---
output: github_document
---

<!-- README.md is generated from README.Rmd. Edit README.Rmd, then run devtools::build_readme(). -->

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

# rddid

<!-- badges: start -->
[![R-CMD-check](https://github.com/dorleventer/rddid/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/dorleventer/rddid/actions/workflows/R-CMD-check.yaml)
[![License: MIT](https://img.shields.io/badge/License-MIT-yellow.svg)](https://opensource.org/licenses/MIT)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

A treatment of interest switches on at a cutoff of a running variable in one
period, the **RD period**. A **confounding policy** switches at the same cutoff,
in every period, so the jump in the outcome at the cutoff in the RD period mixes
the treatment effect with the **confounding jump**. In the **comparison
periods** the treatment of interest is uniform at the cutoff (nobody treated, or
everybody treated), so, provided the treatment of interest has no anticipation or carry-over effects there (which the paper assumes), the jump there *is* the confounding jump. `rddid()`
estimates the jump in every period by local-linear RD and subtracts a weighted
average of the comparison-period jumps from the RD-period jump. How the weights
are set is the **confounding-trend assumption**: constant (equal weights) or
linear in time.

Leventer, D. and D. Nevo (2024). *Correcting Invalid Regression Discontinuity
Designs Using Multiple Time-Period Data.* [arXiv:2408.05847](https://arxiv.org/abs/2408.05847).
`citation("rddid")` returns this reference, with a BibTeX entry.

Compared with a single `rdrobust` fit, `rddid()` runs one local-linear RD per
period with the same conventions (`rd_period()` reproduces `rdrobust` at a given
bandwidth) and then nets the comparison-period jumps out of the RD-period jump,
with standard errors that account for the same units appearing in several
periods and bandwidth rules chosen for the combined estimate.

## Installation

``` r
# install.packages("devtools")
devtools::install_github("dorleventer/rddid")
```

## Quick start

`rddid_sim` is a simulated panel: 1,000 units in three years, a confounding
jump of 0.5 in every year, and a treatment effect of 1 in year 3, the RD period.

```{r quick-start}
library(rddid)
fit <- rddid(rddid_sim, y = "Y", x = "R", time = "year", id = "id", t_rd = 3)
fit
```

The ATT is the effect of the treatment on the units just above the cutoff, the
ones treated in year 3. Years 1 and 2 are the comparison periods, with equal
weights. `h` is the main bandwidth, one for every year, chosen for the RD-DID
estimate; `b` is the pilot bandwidth of the bias correction in each year.
`Conventional` is the local-linear estimate with its conventional standard
error; `Robust (bias-corrected)` is the bias-corrected estimate with its robust
standard error. A plain RD in year 3 would target 1.5, the effect plus the
confounding jump. [Get started](https://dorleventer.github.io/rddid/articles/rddid-estimation.html)
reads every line.

When the running variable moves over time, so that units can change side of the
cutoff between periods, four tests check the assumptions this adds, for example
type continuity:

```{r tests}
rd_typecont(rddid_sim_pv, x = "R", time = "year", id = "id", t_rd = 3)
```

The other three are `rd_compstable()`, `rd_homog()` and `rd_trendcell()`.
`plot()` draws any fit or test (one RD plot per period; the binned type shares
with their fitted lines; the within-type jumps), and `plot_switchers()` shows who
changes side of the cutoff between two periods.

## Learn more

- [Get started](https://dorleventer.github.io/rddid/articles/rddid-estimation.html):
  the estimate, its printout, and the four tests.
- [Checking the identification assumptions](https://dorleventer.github.io/rddid/articles/rddid-validation-tests.html):
  the four tests, one at a time.
- [Plots of the estimate and the checks](https://dorleventer.github.io/rddid/articles/rddid-plots.html):
  `plot()` on a fit and on each test.
- [Reference](https://dorleventer.github.io/rddid/reference/index.html): every
  function, with examples.
