---
output: github_document
---

<!-- README.md is generated from README.Rmd. Edit README.Rmd, then run devtools::build_readme(). -->

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

# rddid

<!-- badges: start -->
[![R-CMD-check](https://github.com/dorleventer/rddid/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/dorleventer/rddid/actions/workflows/R-CMD-check.yaml)
[![License: MIT](https://img.shields.io/badge/License-MIT-yellow.svg)](https://opensource.org/licenses/MIT)
<!-- badges: end -->

Estimation and inference for regression-discontinuity
difference-in-discontinuities (RD-DID) designs, where a confounding policy
switches at the same cutoff as the treatment of interest. Implements the
framework of Leventer and Nevo.

Paper: <https://arxiv.org/abs/2408.05847>

The data are long, one row per unit-period, with an outcome, a running
variable, a period and (for a panel) a unit id. Periods are of two kinds: in
the RD period the treatment of interest switches at the cutoff; in a
comparison period it does not, while the confounding policy switches in every
period. `rddid()` nets the comparison-period discontinuities out of the
RD-period one. When the running variable varies over time within a unit, four
validation tests (`rd_typecont()`, `rd_compstable()`, `rd_homog()`,
`rd_trendcell()`) check the identifying assumptions of Section 4 of the paper.

## Installation

``` r
# install.packages("devtools")
devtools::install_github("dorleventer/rddid")
```

## Quick start

Below, periods 1 and 2 are comparison periods (the confounding discontinuity
is 1 in both) and period 3 is the RD period, where the treatment of interest
adds 1 at the cutoff.

```{r quick-start}
library(rddid)

set.seed(1)
n <- 2000
R <- runif(n, -1, 1)                 # time-invariant running variable, cutoff at 0
u <- rnorm(n, 0, 0.5)                # unit effect
dat <- do.call(rbind, lapply(1:3, function(t) {
  V <- as.integer(R >= 0)            # confounding treatment: sharp RD in every period
  W <- V * (t == 3)                  # treatment of interest: sharp RD in period 3 only
  data.frame(id = seq_len(n), t = t, R = R,
             Y = R + R^2 * (R >= 0) + 1 * V + 1 * W + u + rnorm(n, 0, 0.5))
}))

rddid(dat, y = "Y", x = "R", time = "t", id = "id",
      t_rd = 3, comparisons = c(1, 2), weights = "constant")
```

Set `estimand = "atu"` when the comparison periods are uniformly treated
instead of uniformly untreated (paper, Section 6).

Three vignettes: the estimator (Get started), its bandwidth rules and
sampling schemes, and, for a time-varying running variable, the composition
validation tests: <https://dorleventer.github.io/rddid/articles/index.html>.
