---
title: "Bandwidth rules and sampling schemes"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Bandwidth rules and sampling schemes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
fmt <- function(x, d = 3) formatC(x, format = "f", digits = d)
```

This article covers the `rddid()` arguments that set the bandwidths, the main
bandwidth `h` (for the point estimate) and the pilot bandwidth `b` (for the bias
correction), and the standard error under the three sampling schemes. It uses
`rddid_sim`, a simulated panel with RD period 3, comparison periods 1 and 2 and
a true effect of 1; [Get started](rddid-estimation.html) introduces the data and
the estimator.

```{r}
library(rddid)
```

## Bandwidth rules

`bwselect` picks the rule. Whatever the rule, `fit$bandwidth$h_by_period` and
`fit$bandwidth$b_by_period` hold the bandwidths used in each period, named by
period (stored with the RD period first; the printout and `summary()` list them in
time order).

### The common bandwidth: `bwselect = "joint"` (the default)

```{r}
fit_joint <- rddid(rddid_sim, y = "Y", x = "R", time = "year", id = "id", t_rd = 3)
fit_joint$bandwidth$h_by_period
fit_joint$bandwidth$b_by_period
```

One main bandwidth for every period, chosen to minimize the asymptotic mean
squared error of the RD-DID estimate, not of each period's jump. The bias and
variance of each period's jump are estimated at that period's own CCT pilot
bandwidths, so the rule does not depend on which period is labeled `t_rd`.
The pilot bandwidth of each period keeps that period's CCT ratio of pilot to
main bandwidth. For year 3, from its own CCT pair:

```{r}
y3  <- subset(rddid_sim, year == 3)
bw3 <- rd_bw_cct(y3$Y, y3$R)
fit_joint$bandwidth$h * bw3[["b"]] / bw3[["h"]]
```

which is the year-3 entry of `b_by_period` above.

### Per-period CCT bandwidths: `bwselect = "cct"`

```{r}
fit_cct <- rddid(rddid_sim, y = "Y", x = "R", time = "year", id = "id", t_rd = 3,
                 bwselect = "cct")
fit_cct$bandwidth$h_by_period
```

Each period gets its own MSE-optimal bandwidth from `rdrobust::rdbwselect`
(Calonico, Cattaneo and Titiunik, 2014), computed by `rd_bw_cct()`, ignoring the
other periods. Each bandwidth is optimal for its own jump; the common bandwidth
is optimal for the difference of jumps that `rddid()` reports.

### Period-specific bandwidths for the aggregate: `bwselect = "iter"`

```{r}
fit_iter <- rddid(rddid_sim, y = "Y", x = "R", time = "year", id = "id", t_rd = 3,
                  bwselect = "iter")
fit_iter$bandwidth$h_by_period
```

One bandwidth per period, found by coordinate descent on the asymptotic mean
squared error of the RD-DID estimate. The paper does not use this rule; it is
kept for simulations. `start` sets where the descent starts: `"hstar"` (the
default) starts every period at the common bandwidth, `"cct"` starts each period
at its own CCT bandwidth, and a named vector gives one starting value per period.

```{r}
fit_iter_cct <- rddid(rddid_sim, y = "Y", x = "R", time = "year", id = "id", t_rd = 3,
                      bwselect = "iter", start = "cct")
fit_iter_cct$bandwidth$h_by_period
```

### Fixed bandwidths

```{r}
fit_fixed <- rddid(rddid_sim, y = "Y", x = "R", time = "year", id = "id", t_rd = 3,
                   h = 0.3)
fit_fixed$bandwidth$method
fit_fixed$bandwidth$h_by_period
```

A given `h` is used in every period and bypasses bandwidth selection; `b`
defaults to `h`.

`regularize` (default `TRUE`) adds an `rdrobust`-style regularization term to
the squared bias in the `"joint"` and `"iter"` rules, so that a near-zero
estimated curvature does not blow the bandwidth up; `reg_const` (default
`r eval(formals(rddid)$reg_const)`) is its constant.

### The rules side by side

The four fits above, with the main bandwidth in each year and the two
estimates (standard errors in parentheses):

```{r, echo = FALSE}
estse <- function(f, row) paste0(fmt(f$estimates[row, "est"]), " (", fmt(f$estimates[row, "se"]), ")")
fits <- list(joint = fit_joint, cct = fit_cct, iter = fit_iter, "fixed h = 0.3" = fit_fixed)
tab <- data.frame(
  rule = names(fits),
  t(sapply(fits, function(f) f$bandwidth$h_by_period[c("1", "2", "3")])),
  Conventional = sapply(fits, estse, row = "Conventional"),
  Robust = sapply(fits, estse, row = "Robust"),
  check.names = FALSE, row.names = NULL
)
names(tab)[2:4] <- c("h, year 1", "h, year 2", "h, year 3")
knitr::kable(tab, digits = 3)
```

The rules give different bandwidths, from
`r fmt(min(sapply(fits, function(f) min(f$bandwidth$h_by_period))))` to
`r fmt(max(sapply(fits, function(f) max(f$bandwidth$h_by_period))))`, and, on
these data, conventional estimates from
`r fmt(min(sapply(fits, function(f) f$estimates["Conventional", "est"])))` to
`r fmt(max(sapply(fits, function(f) f$estimates["Conventional", "est"])))`,
with standard errors of about
`r fmt(mean(sapply(fits, function(f) f$estimates["Conventional", "se"])), 2)`;
the truth is 1.

## Sampling schemes

The sampling scheme describes how the data were collected:

- `"cs"`, repeated cross-section: different units in each period;
- `"pc"`, panel, running variable fixed over time: the same units, each on the
  same side of the cutoff in every period;
- `"pv"`, panel, running variable varies over time: some units change side
  between periods (switchers).

With `scheme = "auto"` (the default) `rddid()` reads the scheme off the data:
no `id`, or no unit observed in more than one period, gives `"cs"`; units
observed in more than one period, each on the same side of the cutoff in every
period, give `"pc"`; any unit on different sides in different periods gives
`"pv"`. Without `id`, `rddid()` treats every row as a different unit and says
so; a panel needs `id`.

Three data sets, one per scheme: a repeated cross-section made from `rddid_sim`
by keeping each unit in one year only, `rddid_sim` itself, and `rddid_sim_pv`,
whose running variable drifts between years.

```{r}
sim_cs <- rddid_sim[rddid_sim$id %% 3 == rddid_sim$year %% 3, ]  # each unit in one year
fit_cs <- rddid(sim_cs, y = "Y", x = "R", time = "year", id = "id", t_rd = 3)
fit_pc <- fit_joint
fit_pv <- rddid(rddid_sim_pv, y = "Y", x = "R", time = "year", id = "id", t_rd = 3)
c(cs = fit_cs$scheme, pc = fit_pc$scheme, pv = fit_pv$scheme)
```

`summary()` ends with the robust standard error under each scheme:

```{r}
summary(fit_pv)
```

The same line for the three data sets:

```{r}
se_tab <- rbind(fit_cs$estimates["Robust", c("se_cs", "se_pc", "se_pv")],
                fit_pc$estimates["Robust", c("se_cs", "se_pc", "se_pv")],
                fit_pv$estimates["Robust", c("se_cs", "se_pc", "se_pv")])
dimnames(se_tab) <- list(c("cross-section data", "panel data, R fixed", "panel data, R varies"),
                         c("cross-section s.e.", "panel, R fixed s.e.", "panel, R varies s.e."))
knitr::kable(se_tab, digits = 3)
```

```{r, include = FALSE}
se <- function(f, s) f$estimates["Robust", s]
stopifnot(
  abs(fit_cs$estimates["Conventional", "est"] - 1) < 4 * fit_cs$estimates["Conventional", "se"],
  abs(fit_pc$estimates["Conventional", "est"] - 1) < 4 * fit_pc$estimates["Conventional", "se"],
  abs(fit_pv$estimates["Conventional", "est"] - 1) < 4 * fit_pv$estimates["Conventional", "se"],
  identical(c(fit_cs$scheme, fit_pc$scheme, fit_pv$scheme), c("cs", "pc", "pv")),
  isTRUE(all.equal(se(fit_cs, "se_cs"), se(fit_cs, "se_pc"))),
  isTRUE(all.equal(se(fit_cs, "se_cs"), se(fit_cs, "se_pv"))),
  isTRUE(all.equal(se(fit_pc, "se_pc"), se(fit_pc, "se_pv"))),
  isTRUE(all.equal(fit_joint$bandwidth$h * bw3[["b"]] / bw3[["h"]],
                   unname(fit_joint$bandwidth$b_by_period["3"])))
)
```

- **Cross-section data.** No unit appears in two periods, so the jumps of
  different periods are independent and the three formulas coincide
  (`r fmt(se(fit_cs, "se_cs"))`).
- **Panel data, running variable fixed** (`rddid_sim`). Every unit enters every
  year's fit, on the same side of the cutoff. Its unit effect makes the yearly
  jumps positively correlated, and because the comparison jumps are subtracted,
  that correlation lowers the variance. The panel standard error is
  `r fmt(se(fit_pc, "se_pc"))`; the cross-section one, which ignores the
  correlation, is `r fmt(se(fit_pc, "se_cs"))`. With nobody
  changing side, the formula for a moving running variable reduces to the one
  for a fixed running variable.
- **Panel data, running variable varies** (`rddid_sim_pv`). Units move, so a
  unit can be above the cutoff in one year and below it in another. The
  cross-period covariances are then small (in large samples they vanish
  relative to the variances); here the three standard errors range from
  `r fmt(min(se_tab[3, ]))` to `r fmt(max(se_tab[3, ]))`.

The scheme describes how the data were sampled; the printout reports the
column for that scheme, and the other columns show the other formulas on the
same data. Setting `scheme` replaces the detection. The scheme sets which
standard error is reported and, under the `"joint"` (default) and `"iter"`
bandwidth rules, it also enters the bandwidth: these rules weigh bias against
the asymptotic variance of the estimate, and that variance
has a cross-period covariance term under `"pc"` only (the term is zero under
`"cs"` and negligible in large samples under `"pv"`). On `rddid_sim`:

```{r}
fit_as_cs <- rddid(rddid_sim, y = "Y", x = "R", time = "year", id = "id", t_rd = 3,
                   scheme = "cs")
rbind(pc = c(h = fit_pc$bandwidth$h, coef(fit_pc)),
      cs = c(h = fit_as_cs$bandwidth$h, coef(fit_as_cs)))
```

Treating the panel as a cross-section changes the common bandwidth and so the
estimate. Only with a fixed `h` or `bwselect = "cct"` do the bandwidths not
depend on the scheme; then the scheme leaves the estimate untouched and changes
only the reported standard error.

## References

Calonico, S., M. D. Cattaneo and R. Titiunik (2014). Robust nonparametric
confidence intervals for regression-discontinuity designs. *Econometrica* 82(6),
2295–2326.

Leventer, D. and D. Nevo (2024). Correcting Invalid Regression Discontinuity
Designs Using Multiple Time-Period Data. arXiv:2408.05847.
<https://arxiv.org/abs/2408.05847>
