---
title: "Get started: estimate an RD-DID effect"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get started: estimate an RD-DID effect}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## The setting

A treatment of interest switches on at a cutoff of a running variable in one
period, the **RD period**. A **confounding policy** switches at the same cutoff,
in every period, so the jump in the outcome at the cutoff in the RD period mixes
the treatment effect with the **confounding jump**. In the **comparison
periods** the treatment of interest is uniform at the cutoff (nobody treated, or
everybody treated), so, provided the treatment of interest has no anticipation or carry-over effects there (which the paper assumes), the jump there *is* the confounding jump. `rddid()`
estimates the jump in every period by local-linear RD and subtracts a weighted
average of the comparison-period jumps from the RD-period jump:
$$
\mathrm{ATT}(t_{\mathrm{RD}}) = D_{t_{\mathrm{RD}}} - \sum_t w_t D_t ,
$$
where $t_{\mathrm{RD}}$ is the RD period, $D_t$ is the jump in the outcome at
the cutoff in period $t$, the sum runs over the comparison periods, and $w_t$ is
the weight of comparison period $t$. How the weights are set is the
**confounding-trend assumption**: equal weights if the confounding jump is the
same in every period, a straight-line extrapolation if it moves linearly in time.

## The data

```{r}
library(rddid)
head(rddid_sim)
```

`rddid_sim` is a simulated panel of 1,000 units observed in three years, in
long format: one row per unit and year. `rddid()` uses four columns: the outcome
`Y`, the running variable `R` (cutoff 0), the period `year` and the unit `id`.
The other two record the design: the confounding policy `V` is on above the
cutoff in every year and raises `Y` by 0.5; the treatment of interest `W` is on
above the cutoff in year 3 only and raises `Y` by 1. So year 3 is the RD period,
years 1 and 2 are comparison periods, and **the true effect is 1**.

With your own data, `rddid()` needs the same shape: a data frame in long format,
one row per unit and period, with the outcome, the running variable, the period
and, for a panel, a unit identifier (the panel need not be balanced). The cutoff
is `c` (default 0), and a unit with `x >= c` is above it. The period column is
usually numeric, a year; a character column works with the default constant
trend, while `trend = "linear"` fits its line on the period values and so needs
a numeric column. By default the comparison periods are all periods other than
`t_rd`, in sorted order.

## Estimate

```{r}
fit <- rddid(rddid_sim, y = "Y", x = "R", time = "year", id = "id", t_rd = 3)
fit
```

```{r, include = FALSE}
fmt <- function(x, d = 2) formatC(x, format = "f", digits = d)
fmtp <- function(p) if (p < 0.001) "below 0.001" else fmt(p, 3)
pp <- summary(fit)$per_period
D <- setNames(pp$jump, pp$period)
stopifnot(  # the worked arithmetic below adds up
  isTRUE(all.equal(sum(pp$coef * pp$jump), unname(coef(fit)[["Conventional"]]))),
  isTRUE(all.equal(unname(D[["3"]]), unname(fit$fits[["3"]]$D)))
)
```

Line by line:

- **`RD-DID estimate of the ATT in period 3`**: the estimand and the RD period
  (`t_rd = 3`). The ATT is the effect of the treatment on the units just above
  the cutoff, which are the ones treated in the RD period. If everybody is
  treated in the comparison periods, `estimand = "atu"` labels the same numbers
  as the ATU (see [the ATU section](#atu) below).
- **`Comparison periods: 1, 2 (constant confounding trend; weights 0.5, 0.5)`**:
  by default every period other than `t_rd` is a comparison period. The default
  `trend = "constant"` assumes the confounding jump is the same in every period,
  so each comparison period gets the same weight and their average stands in for
  the confounding jump in year 3.
- **`Sampling scheme`**: how the data were sampled, read off `id` and `R`: here
  a panel (the same units every year) in which no unit changes side of the
  cutoff, because the running variable does not move. The
  standard error takes into account that the same units enter every year's fit.
  Under the default bandwidth rule the scheme can also move the bandwidth, and
  so the estimate; [Bandwidth rules and sampling schemes](rddid-options.html)
  shows how.
- **`Bandwidth: common h`**: the default rule, `bwselect = "joint"`, uses one
  main bandwidth `h` in every period, chosen to minimize the asymptotic mean
  squared error of the RD-DID estimate. **`Pilot bandwidth b (period = value)`**
  lists the bandwidth of the bias correction in each period, in time order.
- **The two rows.** `Conventional` is the local-linear estimate with its
  conventional standard error. `Robust (bias-corrected)` is the bias-corrected
  estimate with its robust standard error (Calonico, Cattaneo and Titiunik,
  2014). Each row has its z statistic, p-value and 95% confidence interval.

The true effect is 1; the two estimates are
`r fmt(coef(fit)[["Conventional"]])` and `r fmt(coef(fit)[["Robust"]])`. A plain
RD in year 3 alone would estimate the effect plus the confounding jump,
1 + 0.5 = 1.5. The year-3 fit is stored in the object:

```{r}
fit$fits[["3"]]
```

Its jump, `r fmt(D[["3"]])`, is that naive RD estimate, whose target is 1.5,
not 1.
`rddid()` subtracts the average of the year-1 and year-2 jumps,
(`r fmt(D[["1"]])` + `r fmt(D[["2"]])`) / 2 = `r fmt((D[["1"]] + D[["2"]]) / 2)`,
its estimate of the confounding jump of 0.5.

## Summary, coefficients and tables

```{r}
summary(fit)
```

`summary()` adds the per-period fits. Each row gives the period's role, its
coefficient in the estimate (+1 for the RD period, minus its weight for a
comparison period), its number of observations, its bandwidths, and its jump
with standard error, conventional (`jump`, `s.e.`) and bias-corrected
(`jump (bc)`, `s.e. (rb)`). The estimate is the sum of coefficient times jump:
`r fmt(D[["3"]], 3)` - 0.5 × `r fmt(D[["1"]], 3)` - 0.5 × `r fmt(D[["2"]], 3)`
= `r fmt(coef(fit)[["Conventional"]], 3)`.

The last lines give the robust standard error under each of the three sampling
schemes. The printout uses the one for the scheme detected in the data; the
other two show what the formulas for the other schemes give on the same data.
Here the cross-section
standard error, `r fmt(fit$estimates["Robust", "se_cs"], 3)`, ignores that the
same units appear in every year; the panel one, which accounts for it, is
`r fmt(fit$estimates["Robust", "se_pc"], 3)`.

```{r}
coef(fit)
confint(fit)
generics::tidy(fit)
```

`coef()` returns the two point estimates and `confint()` their confidence
intervals. `tidy()` (from the generics package) returns one row per estimate, so
table makers such as modelsummary work with `rddid()` fits; [A table for a
paper](#a-table-for-a-paper) below has an example.

## A confounding jump that moves linearly

```{r}
fit_lin <- rddid(rddid_sim, y = "Y", x = "R", time = "year", id = "id", t_rd = 3,
                 trend = "linear")
fit_lin
```

With `trend = "linear"` the confounding jump may move linearly in time. The line
through the year-1 and year-2 jumps is extrapolated to year 3, which gives
weights -1 and 2: the confounding jump in year 3 is estimated by
$2 D_2 - D_1$. This needs at least two comparison periods. The standard error
grows (`r fmt(fit_lin$estimates["Conventional", "se"], 3)` against
`r fmt(fit$estimates["Conventional", "se"], 3)`) because extrapolating a line
amplifies noise. In `rddid_sim` the confounding jump is constant, so both
estimates are valid; the linear one gives up precision to allow for a linear
trend. A numeric vector instead of `"constant"` or `"linear"` sets the weights
directly, one per comparison period.

## A table for a paper

`tidy()` gives the estimates and `glance()` the design (RD period, comparison
periods, trend, bandwidth rule, `h`, scheme), which is what table makers read.
With the modelsummary package installed, the two fits side by side, with
confidence intervals:

```{r modelsummary, eval = requireNamespace("modelsummary", quietly = TRUE) && requireNamespace("broom", quietly = TRUE)}
modelsummary::modelsummary(
  list(Constant = fit, Linear = fit_lin), statistic = "conf.int",
  gof_map = list(list(raw = "h", clean = "h", fmt = 3),
                 list(raw = "nobs", clean = "Num.Obs.", fmt = 0)))
```

The same call with `output = "rddid_table.tex"` (or `.html`, `.md`, `.docx`)
writes the table to a file; `write.csv(generics::tidy(fit), "rddid_estimate.csv",
row.names = FALSE)` writes the estimate table to a plain CSV file.

`gof_map` sets which rows of `glance()` appear below the estimates and their
digits (`fmt`): here `h` with three digits and the number of observations.
Without it, every column of `glance()` is listed and `h` prints with all its
digits. The last line writes the estimate table to a plain CSV file.

## When the running variable moves over time

In `rddid_sim` every unit stays on its side of the cutoff. In many applications
the running variable moves (a population count, a test score, an income), and
some units, the **switchers**, are above the cutoff in one period and below it
in another. The units at the cutoff in the RD period can then be a different mix
from the units at the cutoff in a comparison period, and the difference of jumps
can be biased. A unit's **type** records its side of the cutoff in periods other
than the one at hand. The tests use two versions:

| Test | A unit's type in period *t* | Printed as |
|---|---|---|
| type continuity, composition stability | its sides of the cutoff in all the other periods, in time order | `++`, `+-`, `-+`, `--` (`+` above, `-` below) |
| homogeneous confounding, constant within-type confounding | its side of the cutoff in the RD period (the default, `type_by = "rd_side"`) | `+`, `-`; in plots "Above in 3", "Below in 3" |

With two periods, both versions are the unit's side of the cutoff in the other
period.

The RD-DID estimate identifies the ATT when the type distribution is
continuous at the cutoff and the confounding jump within each type is constant
over time, and in addition either the type composition is stable across periods
or the confounding jump is the same for every type (Leventer and Nevo, 2024).
Four tests check these assumptions. `rddid_sim_pv` is the `rddid_sim` design
with a running variable that drifts between years. The estimate whose
assumptions the four tests below check:

```{r}
fit_pv <- rddid(rddid_sim_pv, y = "Y", x = "R", time = "year", id = "id", t_rd = 3)
fit_pv
```

The sampling scheme is now a panel in which some units change side of the
cutoff between years.

**Type continuity.** H0: the share of each type jumps by zero at the cutoff, in
every period. A rejection means units sort around the cutoff by their side in
other periods, so the jump in the outcome partly reflects who the units are; the
estimate can then be biased, whatever the other tests say.

```{r}
tc <- rd_typecont(rddid_sim_pv, x = "R", time = "year", id = "id", t_rd = 3)
tc
```

Every test prints its null (H0) and the sampling scheme read off the data.
`Types: ++, +-, -+, --` are the types of the first row of the table above.
`Bandwidth: CCT MSE-optimal, chosen per cell` means that each local-linear fit
the test runs (one per period and type, a *cell*) gets its own MSE-optimal
bandwidth from `rdrobust::rdbwselect` (Calonico, Cattaneo and Titiunik, 2014;
hence "CCT"). `Joint Wald chi-squared(9)` tests all the jumps in the type shares
at once: in each year the four shares sum to one, so three jumps are free, and
three years give 9. Each `Period` line tests one year alone.

**Composition stability.** H0: the share of each type among the units just above
the cutoff is the same in the RD period and in each comparison period. A rejection
means the units at the cutoff in the RD period are a different mix of types from
those in a comparison period; the estimate is then valid only if the next test's
null holds.

```{r}
cs <- rd_compstable(rddid_sim_pv, x = "R", time = "year", id = "id", t_rd = 3)
cs
```

`Pair 3::1` compares year 3, the RD period, with year 1; its `chi-squared(3)`
tests the three free jumps in the type shares. `n above the cutoff` counts the
units above the cutoff in the RD year, in the comparison year, and in both.
`Joint over pairs` adds up the pair statistics and their degrees of freedom; it
treats the pairs as independent although they share the year-3 units, so its
p-value is approximate (the paper's test is per pair).

**Homogeneous confounding.** H0: in each comparison period the confounding jump
is the same for every type. Here, and in the next test, the types are those of
the second row of the table above: a unit's side of the cutoff in the RD period.
A rejection means the confounding jump differs by type; together with a
rejection of composition stability, the comparison-period jumps do not measure
the confounding jump of the units at the RD-period cutoff, and the estimate can
be biased.

```{r}
hg <- rd_homog(rddid_sim_pv, y = "Y", x = "R", time = "year", id = "id", t_rd = 3)
hg
```

The table gives the confounding jump of each type in each comparison year, with
its standard error and number of observations. The Wald statistic has one
contrast per year, type `+` against the type marked `(reference)`, so two
comparison years give `chi-squared(2)`.

**Constant within-type confounding.** H0: within each type, the confounding jump
is the same in every comparison period. This is a pre-trends check: the
assumption concerns the RD period, where the confounding jump cannot be seen
apart from the effect, so the test asks whether it is stable across the
comparison periods. A rejection means the constant-trend weights do not cancel
the confounding jump. `trend = "linear"` allows a confounding jump that moves
linearly in time; `rd_trendcell(trend = "linear")` tests that version (three or
more comparison periods).

```{r}
tr <- rd_trendcell(rddid_sim_pv, y = "Y", x = "R", time = "year", id = "id", t_rd = 3)
tr
```

The jumps are those of `rd_homog()`, grouped by type. The Wald statistic has
one contrast per type: year 2 against the year marked `(reference)`, the type's
first comparison year.

One table for all four:

```{r}
do.call(rbind, lapply(list(tc, cs, hg, tr), generics::tidy))
```

The composition-stability row is the joint-over-pairs statistic, whose p-value
is approximate; the per-pair tests are in its printout above.

In `rddid_sim_pv` three of the four nulls are true by construction: types are
continuous at the cutoff, and the confounding jump is 0.5 for every unit in
every year. Composition stability is false: units at the year-3 cutoff were
mostly on the same side in years 1 and 2, while units at the year-1 cutoff are
spread evenly over the four types. In this sample the four p-values are
`r fmtp(tc$p_value)`, `r fmtp(cs$p_value)`, `r fmtp(hg$p_value)` and
`r fmtp(tr$p_value)` (in the order of the table). A p-value below 0.05 for a true null is a false
rejection, which a test at the 5% level is designed to make in about one sample in twenty.
Because the other three assumptions hold, homogeneous confounding among them,
the failure of composition stability does not bias `fit_pv`: its estimate is `r fmt(coef(fit_pv)[["Conventional"]])`,
against a true effect of 1. The article on the identification assumptions shows
how each statistic is built and the options of the tests.

## Comparison periods where everybody is treated: the ATU {#atu}

When everybody is treated in the comparison periods, rather than nobody, the
same difference of jumps estimates the ATU: the effect for the units just below
the cutoff, which are untreated in the RD period. `estimand = "atu"` labels the
output as the ATU: the estimate, standard errors and bandwidths are the same as
without it (among the tests, it changes only `rd_compstable()`). For
illustration, make everybody treated in years 1 and 2 of `rddid_sim`, with the
same effect of 1:

```{r}
sim_atu <- transform(rddid_sim, W = ifelse(year < 3, 1, W), Y = Y + (year < 3))
rddid(sim_atu, y = "Y", x = "R", time = "year", id = "id", t_rd = 3, estimand = "atu")
```

The numbers are those of `fit`, now labeled the ATU; the true ATU here is 1.

## What the paper reports

- **The estimate.** The paper reports both rows. It bases its confidence
  intervals and p-values on the `Robust (bias-corrected)` row, following
  Calonico, Cattaneo and Titiunik (2014): at a bandwidth chosen to minimize the
  mean squared error, the conventional interval is centered on a biased estimate
  and can under-cover.
- **The design.** The choices behind an estimate (the RD period, the comparison
  periods, the trend assumption with its weights, the bandwidth rule and `h`,
  the sampling scheme) are in the first lines of the printout;
  `generics::glance(fit)` returns them as a one-row table.
- **The tests.** When the running variable moves over time, the table above
  collects the four tests of the assumptions the estimate relies on.

## Next

- [Checking the identification assumptions](rddid-validation-tests.html): how
  each test statistic is built, the tests under the ATU design, and their
  options.
- [Plots of the estimate and the checks](rddid-plots.html): `plot()` on a fit
  and on each test, the switchers picture, and saving a figure.
- [Bandwidth rules and sampling schemes](rddid-options.html): `bwselect`, fixed
  bandwidths, and the standard error under each scheme.
- [How rddid() computes the estimate](rddid-how-it-works.html): the estimate and
  the tests rebuilt by hand, for referees and readers who want to see the
  pieces.

## References

`citation("rddid")` returns the reference for the package, the paper below, with
a BibTeX entry.

Leventer, D. and D. Nevo (2024). Correcting Invalid Regression Discontinuity
Designs Using Multiple Time-Period Data. arXiv:2408.05847.
<https://arxiv.org/abs/2408.05847>

Calonico, S., M. D. Cattaneo and R. Titiunik (2014). Robust nonparametric
confidence intervals for regression-discontinuity designs. *Econometrica* 82(6),
2295–2326.
