---
title: "Composition validation tests"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Composition validation tests}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 3)
```

```{r setup, message = FALSE}
library(rddid)
library(ggplot2)
fmt <- function(x, d = 3) formatC(x, format = "f", digits = d)
fmtp <- function(p) ifelse(p < 0.001, "< 0.001", fmt(p))          # p-values in tables
pmath <- function(p) ifelse(p < 0.001, "p < 0.001", paste("p =", fmt(p)))  # p-values in prose
bin_x <- function(x, w = 0.1) ifelse(x >= 0, floor(x / w) * w + w / 2, -(floor(-x / w) * w + w / 2))
```

## Setup

The four tests in this vignette apply when the running variable $R$ varies over time within a unit (the `pv` sampling scheme); with a time-constant running variable every unit is on the same side of the cutoff in every period and the objects below are degenerate. When $R$ varies, a unit can sit on different sides of the cutoff in different periods. Its **type** in period $t$ is its side of the cutoff in the *other* period(s); $\pi_{t,(+)}(v)$ and $\pi_{t,(-)}(v)$ are the shares of type $v$ just above / below the cutoff in period $t$, and $\alpha_{t,0}(v)$ is the within-type confounding discontinuity in period $t$. Section 4 of the paper states four assumptions on these objects and Section 4.4 gives a test for each. Every test function below takes the same long data frame (`id`, `t`, `R`, `Y`) that `rddid()` takes.

```{r dgp}
m_b <- function(r) r + r^2 * (r >= 0)
dgp_b <- function(n = 4000, d = 0, alpha = c(1, 1), kappa = 0, sort_p = 0, sort_w = 0.5,
                  tau = 1, periods = 2, gamma = 0, seed = 1) {
  set.seed(seed)
  eta <- rnorm(n)                                          # latent level of the running variable
  R <- sapply(seq_len(periods), function(t) eta + d * (t - 1) + rnorm(n, 0, 0.5))
  if (sort_p > 0) {                                        # sorting: some units just below the RD-period
    V1 <- R[, 1] >= 0; r2 <- R[, periods]                  # cutoff that were above in period 1 move above
    mover <- r2 > -sort_w & r2 < 0 & V1 & runif(n) < sort_p
    R[mover, periods] <- -R[mover, periods]
  }
  V <- R >= 0
  do.call(rbind, lapply(seq_len(periods), function(t) {
    other <- if (periods == 2) V[, 3 - t] else V[, periods]  # the unit's type: side in the other period
    a <- alpha[other + 1] + gamma * t                        # within-type confounding discontinuity
    data.frame(id = seq_len(n), t = t, R = R[, t],
               Y = m_b(R[, t]) + a * V[, t] + tau * V[, t] * (t == periods) + kappa * eta + rnorm(n, 0, 0.5))
  }))
}
```

- `d`: common drift of $R$ between periods — breaks composition stability.
- `alpha`: the confounding jump by type, `c(alpha(0), alpha(1))` — unequal values break homogeneous confounding.
- `kappa`, `sort_p`: sorting of period-1-above units across the period-2 cutoff, with `kappa` making the outcome level depend on the type — breaks continuity of the type distribution.
- `gamma`: a trend in the confounding across periods (three-period version) — breaks constant confounding within types.

```{r scenarios}
S0   <- dgp_b()                                # every assumption holds
S1   <- dgp_b(d = 0.5)                         # composition changes; confounding homogeneous
S2   <- dgp_b(d = 0.5, alpha = c(0.5, 1.5))    # composition changes; confounding heterogeneous
S3   <- dgp_b(kappa = 2, sort_p = 0.5)         # sorting at the RD-period cutoff
S0_3 <- dgp_b(periods = 3)                     # three periods, everything holds
S4   <- dgp_b(periods = 3, gamma = 0.3)        # confounding trends across periods
```

```{r type-pattern-plot, fig.alt = "Scatter of R in period 1 vs period 2 for S1, coloured by side pattern"}
w1 <- reshape(S1[, c("id", "t", "R")], idvar = "id", timevar = "t", direction = "wide")
names(w1) <- c("id", "R1", "R2")
w1$pattern <- with(w1, paste0(ifelse(R1 >= 0, "+", "-"), ifelse(R2 >= 0, "+", "-")))
ggplot(w1, aes(R1, R2, colour = pattern)) +
  geom_point(size = 0.5, alpha = 0.4) +
  geom_hline(yintercept = 0, linetype = "dashed") + geom_vline(xintercept = 0, linetype = "dashed") +
  labs(colour = "sign(R1)sign(R2)") + theme_bw()
```

With drift (`d = 0.5`), units just above the period-2 cutoff were mostly below the cutoff in period 1.

## Continuous type distribution — `rd_typecont()`

Continuous type distribution requires $\pi_{t,(+)}(v) = \pi_{t,(-)}(v)$ in every period. The test regresses, in each period $t$, the indicator of the unit's side in the other period $s$, $1\{V_{i,s} = 1\}$, on $R_{i,t}$: the local-linear RD jump of that indicator is $\hat\pi_{t,(+)}(1) - \hat\pi_{t,(-)}(1)$. Building the two per-period jumps directly with `rd_bw_cct()` + `rd_period()`:

```{r typecont-jumps}
side_share <- function(dat) {
  w <- reshape(dat[, c("id", "t", "R")], idvar = "id", timevar = "t", direction = "wide")
  names(w) <- c("id", "R1", "R2")
  V <- w[, c("R1", "R2")] >= 0
  lapply(1:2, function(t) {
    other <- V[, 3 - t]
    d <- data.frame(id = w$id, R = w[[paste0("R", t)]], ind = as.integer(other))
    bw <- rd_bw_cct(d$ind, d$R, c = 0)
    fit <- rd_period(d$ind, d$R, h = unname(bw["h"]), b = unname(bw["b"]), id = d$id, c = 0)
    list(d = d, fit = fit)
  })
}
ss0 <- side_share(S0); ss3 <- side_share(S3)
tab_ss <- data.frame(
  scenario = rep(c("S0", "S3"), each = 2), t = rep(1:2, 2),
  jump = c(sapply(ss0, function(s) s$fit$D), sapply(ss3, function(s) s$fit$D)),
  se   = c(sapply(ss0, function(s) sqrt(s$fit$V_D)), sapply(ss3, function(s) sqrt(s$fit$V_D)))
)
knitr::kable(tab_ss, digits = 3, col.names = c("scenario", "t", "jump", "SE"))
```

`rd_typecont()` computes the same jumps and combines them into a joint Wald statistic across periods, with the cross-period covariance of the two jumps (the same units enter both regressions) inside it. The statistic, its degrees of freedom and p-value are `$ll_wald$stat`, `$df`, `$p`; the per-period versions sit in `$per_period[[t]]$ll_wald`:

```{r typecont-wald}
tc0 <- rd_typecont(S0, x = "R", time = "t", id = "id", bwselect = "cct", bc = FALSE)
tc3 <- rd_typecont(S3, x = "R", time = "t", id = "id", bwselect = "cct", bc = FALSE)
tc0_bc <- rd_typecont(S0, x = "R", time = "t", id = "id", bwselect = "cct", bc = TRUE)
```

For S0, $\chi^2(`r tc0$ll_wald$df`) = `r fmt(tc0$ll_wald$stat)`$, $`r pmath(tc0$ll_wald$p)`$ (per period $`r pmath(tc0$per_period[["1"]]$ll_wald$p)`$ and $`r pmath(tc0$per_period[["2"]]$ll_wald$p)`$); with `bc = TRUE` the Wald uses the bias-corrected jumps and their robust variance, $`r pmath(tc0_bc$ll_wald$p)`$. For S3, $\chi^2(`r tc3$ll_wald$df`) = `r fmt(tc3$ll_wald$stat)`$, $`r pmath(tc3$ll_wald$p)`$ (per period $`r pmath(tc3$per_period[["1"]]$ll_wald$p)`$ and $`r pmath(tc3$per_period[["2"]]$ll_wald$p)`$). S0's per-period jumps are indistinguishable from zero in either period; S3's are not, in either period.

```{r typecont-plot, fig.height = 3.5, fig.alt = "Type-share RD plots, rows S0/S3, columns period 1/2"}
type_share_df <- function(ss, scenario) {
  do.call(rbind, lapply(1:2, function(t) {
    fit <- ss[[t]]$fit; d <- ss[[t]]$d
    agg <- aggregate(d$ind, by = list(R = bin_x(d$R)), FUN = mean)
    names(agg)[2] <- "ind"; agg$scenario <- scenario; agg$t <- t
    h <- unname(fit$h)
    lines <- rbind(
      data.frame(R = c(-h, 0), ind = fit$sides[["-"]]$beta0 + fit$sides[["-"]]$slope * (c(-h, 0) - 0), side = "-"),
      data.frame(R = c(0, h),  ind = fit$sides[["+"]]$beta0 + fit$sides[["+"]]$slope * (c(0, h) - 0),  side = "+")
    )
    lines$scenario <- scenario; lines$t <- t
    list(binned = agg, lines = lines)
  }))
}
b0 <- type_share_df(ss0, "S0"); b3 <- type_share_df(ss3, "S3")
binned <- rbind(do.call(rbind, b0[, "binned"]), do.call(rbind, b3[, "binned"]))
lines  <- rbind(do.call(rbind, b0[, "lines"]),  do.call(rbind, b3[, "lines"]))
ggplot(binned, aes(R, ind)) +
  geom_point(size = 0.9, colour = "grey45") +
  geom_line(data = lines, aes(group = side), colour = "firebrick", linewidth = 1.1) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  facet_grid(scenario ~ t, labeller = label_both) + labs(y = "other-period type indicator") + theme_bw()
```

## Composition stability — `rd_compstable()`

Composition stability requires $\pi_{t_{\mathrm{RD}},(+)}(v) = \pi_{t_0,(+)}(v)$. The test reflects the comparison-period above-cutoff units onto a negative axis and stacks them with the RD-period above-cutoff units: for S1 ($t_{\mathrm{RD}} = 2$, $t_0 = 1$),

```{r compstable-build}
w1 <- reshape(S1[, c("id", "t", "R")], idvar = "id", timevar = "t", direction = "wide")
names(w1) <- c("id", "R1", "R2")
above1 <- w1$R1 >= 0; above2 <- w1$R2 >= 0
refl <- rbind(
  data.frame(id = w1$id[above1], x = -(w1$R1[above1]), ind = as.integer(above2[above1])),
  data.frame(id = w1$id[above2], x =   w1$R2[above2],  ind = as.integer(above1[above2]))
)
bw_r <- rd_bw_cct(refl$ind, refl$x, c = 0)
fit_r <- rd_period(refl$ind, refl$x, h = unname(bw_r["h"]), b = unname(bw_r["b"]), id = refl$id, c = 0)
```

```{r compstable-plot, fig.alt = "Binned means and LL line of the reflected composition-stability RD, S1"}
agg_r <- aggregate(refl$ind, by = list(x = bin_x(refl$x)), FUN = mean)
names(agg_r)[2] <- "ind"
h_r <- unname(fit_r$h)
lines_r <- rbind(
  data.frame(x = c(-h_r, 0), ind = fit_r$sides[["-"]]$beta0 + fit_r$sides[["-"]]$slope * (c(-h_r, 0) - 0), side = "-"),
  data.frame(x = c(0, h_r),  ind = fit_r$sides[["+"]]$beta0 + fit_r$sides[["+"]]$slope * (c(0, h_r) - 0),  side = "+")
)
ggplot(agg_r, aes(x, ind)) +
  geom_point(size = 0.9, colour = "grey45") + geom_line(data = lines_r, aes(group = side), colour = "firebrick", linewidth = 1.1) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  labs(x = "reflected running variable", y = "other-side type indicator") + theme_bw()
```

The jump at $x = 0$ estimates $\pi_{2,(+)}(1) - \pi_{1,(+)}(1)$. `rd_compstable()` runs the same construction and reports the jump, its SE, and the Wald test:

```{r compstable-wald}
cs0 <- rd_compstable(S0, x = "R", time = "t", id = "id", t_rd = 2, comparisons = 1, bwselect = "cct", bc = FALSE)
cs1 <- rd_compstable(S1, x = "R", time = "t", id = "id", t_rd = 2, comparisons = 1, bwselect = "cct", bc = FALSE)
p1 <- cs1$pairs[["2::1"]]; p0 <- cs0$pairs[["2::1"]]
```

For S1: jump $= `r fmt(p1$jumps)`$, SE $= `r fmt(p1$jump_se)`$, $\chi^2(`r p1$ll_wald$df`) = `r fmt(p1$ll_wald$stat)`$, $`r pmath(p1$ll_wald$p)`$, with `r p1$n_trd` above-cutoff units from $t_{\mathrm{RD}}$, `r p1$n_t0` from $t_0$, and `r p1$n_both` above the cutoff in both periods. The package SE accounts for those `r p1$n_both` doubly-counted units through the id-matched cross-side covariance term. For S0, for contrast: jump $= `r fmt(p0$jumps)`$, $`r pmath(p0$ll_wald$p)`$.

## ATU designs — estimand = "atu"

With comparison periods uniformly treated, the ATU requires composition
stability of the below-cutoff shares,
$\pi_{t_{\mathrm{RD}},(-)}(v) = \pi_{t_0,(-)}(v)$ (paper, Section 6);
`rd_compstable(estimand = "atu")` mirrors the running variable and runs the
same construction on the below-cutoff units.

```{r compstable-atu}
cs1_atu <- rd_compstable(S1, x = "R", time = "t", id = "id", t_rd = 2, comparisons = 1,
                         bwselect = "cct", bc = FALSE, estimand = "atu")
p1_atu <- cs1_atu$pairs[["2::1"]]
```

For S1, the `"att"` call above gives jump $= `r fmt(p1$jumps)`$,
$\chi^2(`r p1$ll_wald$df`) = `r fmt(p1$ll_wald$stat)`$, $`r pmath(p1$ll_wald$p)`$,
on `r p1$n_trd` above-cutoff units from $t_{\mathrm{RD}}$ and `r p1$n_t0` from
$t_0$; the `"atu"` call gives jump $= `r fmt(p1_atu$jumps)`$,
$\chi^2(`r p1_atu$ll_wald$df`) = `r fmt(p1_atu$ll_wald$stat)`$,
$`r pmath(p1_atu$ll_wald$p)`$, on `r p1_atu$n_trd` below-cutoff units from
$t_{\mathrm{RD}}$ and `r p1_atu$n_t0` from $t_0$ -- different units from the
`"att"` call. Both reject under S1's drift.

The other three tests are unchanged under `estimand = "atu"`:

```{r typecont-atu}
tc_att <- rd_typecont(S1, x = "R", time = "t", id = "id", bwselect = "cct", bc = FALSE)
tc_atu <- rd_typecont(S1, x = "R", time = "t", id = "id", bwselect = "cct", bc = FALSE,
                      estimand = "atu")
```

`rd_typecont()` on S1 gives the same joint statistic either way,
$\chi^2(`r tc_att$ll_wald$df`) = `r fmt(tc_att$ll_wald$stat)`$ under `"att"`
and `r fmt(tc_atu$ll_wald$stat)` under `"atu"`.

```{r checks-atu-section, include = FALSE}
stat_att <- tc_att$ll_wald$stat
stat_atu <- tc_atu$ll_wald$stat
stopifnot(
  isTRUE(all.equal(stat_att, stat_atu)),
  !isTRUE(all.equal(p1$ll_wald$stat, p1_atu$ll_wald$stat))
)
```

## Homogeneous confounding — `rd_homog()`

Homogeneous confounding requires $\alpha_{t_0,0}(0) = \alpha_{t_0,0}(1)$ in the comparison period. The test estimates the outcome RD jump within each type (the unit's side of the cutoff in the RD period, `type_by = "rd_side"`, the default), in the comparison period, and tests equality:

```{r homog}
h1 <- rd_homog(S1, y = "Y", x = "R", time = "t", id = "id", comparisons = 1, t_rd = 2, bwselect = "cct", bc = FALSE)
h2 <- rd_homog(S2, y = "Y", x = "R", time = "t", id = "id", comparisons = 1, t_rd = 2, bwselect = "cct", bc = FALSE)
tab_h <- rbind(cbind(scenario = "S1", h1$period_type_jumps), cbind(scenario = "S2", h2$period_type_jumps))
knitr::kable(tab_h[, c("scenario", "period", "type", "jump", "se", "n")], digits = 3, row.names = FALSE)
```

S1: $\chi^2(`r h1$df`) = `r fmt(h1$statistic)`$, $`r pmath(h1$p_value)`$. S2: $\chi^2(`r h2$df`) = `r fmt(h2$statistic)`$, $`r pmath(h2$p_value)`$. In S1 the two within-type comparison-period jumps, $`r fmt(h1$period_type_jumps$jump[1])`$ and $`r fmt(h1$period_type_jumps$jump[2])`$, are close to each other (both near $\alpha = 1$); in S2 they are $`r fmt(h2$period_type_jumps$jump[1])`$ and $`r fmt(h2$period_type_jumps$jump[2])`$, close to the construction values `alpha = c(0.5, 1.5)`.

```{r homog-plot, fig.alt = "Point-range of within-type jumps, facets S1/S2"}
tab_h$scenario <- factor(tab_h$scenario, levels = c("S1", "S2"))
ggplot(tab_h, aes(type, jump)) +
  geom_pointrange(aes(ymin = jump - 1.96 * se, ymax = jump + 1.96 * se)) +
  facet_wrap(~ scenario) + labs(x = "type (RD-period side)", y = "within-type jump") + theme_bw()
```

## Constant confounding discontinuity within types — `rd_trendcell()`

Constant confounding within types requires $\alpha_{t_{\mathrm{RD}},0}(v) = \alpha_{t_0,0}(v)$: the same within-type confounding jump at the RD period as at the comparison period. It is not directly testable at the RD period; with two comparison periods ($t = 1, 2$, RD period $t = 3$) the test instead checks whether the within-type jumps are constant *across* the two comparison periods, a pre-trends check in the difference-in-differences sense:

```{r trendcell}
tr0 <- rd_trendcell(S0_3, y = "Y", x = "R", time = "t", id = "id", comparisons = c(1, 2), t_rd = 3, bwselect = "cct", trend = "constant", bc = FALSE)
tr4 <- rd_trendcell(S4, y = "Y", x = "R", time = "t", id = "id", comparisons = c(1, 2), t_rd = 3, bwselect = "cct", trend = "constant", bc = FALSE)
tab_tr <- rbind(cbind(scenario = "S0_3", tr0$cell_period_jumps), cbind(scenario = "S4", tr4$cell_period_jumps))
knitr::kable(tab_tr[, c("scenario", "cell", "period", "jump", "se", "n")], digits = 3, row.names = FALSE)
```

S0_3: $\chi^2(`r tr0$df`) = `r fmt(tr0$statistic)`$, $`r pmath(tr0$p_value)`$. S4: $\chi^2(`r tr4$df`) = `r fmt(tr4$statistic)`$, $`r pmath(tr4$p_value)`$. In S4 each cell's jump rises from $t = 1$ to $t = 2$: cell "+" from $`r fmt(tr4$cell_period_jumps$jump[1])`$ to $`r fmt(tr4$cell_period_jumps$jump[2])`$, cell "-" from $`r fmt(tr4$cell_period_jumps$jump[3])`$ to $`r fmt(tr4$cell_period_jumps$jump[4])`$, by about the construction increment `gamma = 0.3`. With only two comparison periods the constant form is what can be tested here; `trend = "linear"` needs three or more.

```{r trendcell-plot, fig.alt = "Point-range of within-cell jumps by period, facets S0_3/S4"}
tab_tr$scenario <- factor(tab_tr$scenario, levels = c("S0_3", "S4"))
ggplot(tab_tr, aes(period, jump, colour = cell)) +
  geom_pointrange(aes(ymin = jump - 1.96 * se, ymax = jump + 1.96 * se), position = position_dodge(width = 0.3)) +
  facet_wrap(~ scenario) + theme_bw()
```

## Summary

```{r summary-tests}
tc1 <- rd_typecont(S1, x = "R", time = "t", id = "id", bwselect = "cct", bc = FALSE)
tc2 <- rd_typecont(S2, x = "R", time = "t", id = "id", bwselect = "cct", bc = FALSE)
cs2 <- rd_compstable(S2, x = "R", time = "t", id = "id", t_rd = 2, comparisons = 1, bwselect = "cct", bc = FALSE)
cs3 <- rd_compstable(S3, x = "R", time = "t", id = "id", t_rd = 2, comparisons = 1, bwselect = "cct", bc = FALSE)
h0 <- rd_homog(S0, y = "Y", x = "R", time = "t", id = "id", comparisons = 1, t_rd = 2, bwselect = "cct", bc = FALSE)
h3 <- rd_homog(S3, y = "Y", x = "R", time = "t", id = "id", comparisons = 1, t_rd = 2, bwselect = "cct", bc = FALSE)

rid <- function(dat, ...) rddid(dat, y = "Y", x = "R", time = "t", id = "id", t_rd = max(dat$t), bwselect = "cct", ...)
estfmt <- function(r) sprintf("%s (%s)", fmt(r$estimates["Conventional", "est"]), fmt(r$estimates["Conventional", "se"]))
r0 <- rid(S0); r1 <- rid(S1); r2 <- rid(S2); r3 <- rid(S3)
r0_3 <- rid(S0_3); r4c <- rid(S4); r0_3l <- rid(S0_3, weights = "linear"); r4l <- rid(S4, weights = "linear")

tab1 <- data.frame(
  scenario = c("S0", "S1", "S2", "S3"),
  type_cont_p    = fmtp(c(tc0$ll_wald$p, tc1$ll_wald$p, tc2$ll_wald$p, tc3$ll_wald$p)),
  comp_stable_p  = fmtp(c(p0$ll_wald$p, p1$ll_wald$p, cs2$pairs[["2::1"]]$ll_wald$p, cs3$pairs[["2::1"]]$ll_wald$p)),
  homog_p        = fmtp(c(h0$p_value, h1$p_value, h2$p_value, h3$p_value)),
  rddid_est_se   = c(estfmt(r0), estfmt(r1), estfmt(r2), estfmt(r3)),
  truth = 1
)
knitr::kable(tab1, col.names = c("scenario", "type-cont p", "comp-stable p", "homog p", "rddid() est (SE)", "truth"))

tab2 <- data.frame(
  scenario = c("S0_3", "S4"), trendcell_p = fmtp(c(tr0$p_value, tr4$p_value)),
  est_constant = c(estfmt(r0_3), estfmt(r4c)), est_linear = c(estfmt(r0_3l), estfmt(r4l)), truth = 1
)
knitr::kable(tab2, col.names = c("scenario", "trend-cell p", "rddid() est (SE), constant", "rddid() est (SE), linear", "truth"))
```

In S0 all four assumptions hold and the estimate, `r estfmt(r0)`, sits at the truth of 1. In S1 composition stability is rejected ($`r pmath(p1$ll_wald$p)`$) while homogeneous confounding is not ($`r pmath(h1$p_value)`$), and the estimate, `r estfmt(r1)`, is still at the truth — the identification result holds under either composition stability or homogeneous confounding. In S2 both are rejected ($`r pmath(cs2$pairs[["2::1"]]$ll_wald$p)`$ and $`r pmath(h2$p_value)`$), and the estimate, `r estfmt(r2)`, is off the truth by the composition term. In S3 type continuity is rejected ($`r pmath(tc3$ll_wald$p)`$) and the estimate, `r estfmt(r3)`, is off. In the three-period design, S4's trend-cell test rejects ($`r pmath(tr4$p_value)`$): the constant-weight estimate, `r estfmt(r4c)`, is off the truth while the linear-weight estimate, `r estfmt(r4l)`, is close to it.

```{r checks, include = FALSE}
stopifnot(
  cs1$pairs[["2::1"]]$ll_wald$p < 0.01, tc3$ll_wald$p < 0.01,
  tc3$per_period[["1"]]$ll_wald$p < 0.01, tc3$per_period[["2"]]$ll_wald$p < 0.01,
  h2$p_value < 0.01, tr4$p_value < 0.05,
  tc0$ll_wald$p > 0.2, tc1$ll_wald$p > 0.2, p0$ll_wald$p > 0.2,
  h1$p_value > 0.2, h0$p_value > 0.2, cs3$pairs[["2::1"]]$ll_wald$p > 0.2, tr0$p_value > 0.2,
  abs(r0$estimates["Conventional", "est"] - 1) < 4 * r0$estimates["Conventional", "se"],
  abs(r1$estimates["Conventional", "est"] - 1) < 4 * r1$estimates["Conventional", "se"],
  abs(r2$estimates["Conventional", "est"] - 1) > 3 * r2$estimates["Conventional", "se"],
  abs(r4c$estimates["Conventional", "est"] - 1) > 3 * r4c$estimates["Conventional", "se"]
)
```
