---
title: "rddid-linear-trend"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{rddid-linear-trend}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Intro

This is an example notebook of how to estimate treatment effects in RDDID designs with a assumption that potential outcome discontinuities evolve linearly over time. The identification framework is based on Corollary 1(ii) in Leventer and Nevo (2024). The estimation framework is presented in Supplementary Appendix.

First, lets install the rddid package.

```{r setup}
library(rddid)
library(tidyverse)
```

# Setup

To create an easy example that can be replicated, we use the simulated data in the paper. 

Lets create a dataset with 3 time periods, with a discontinuity of 100 in time 1, 200 in time 2, and 400 in time 3 (assuming time 3 is the RD period of interest). If the discontinuity evolves linearly, then it should be 300 in time 3 without the RD, implying the treatment effect is $400-300=100$.

We'll manipulate the simulated data (provided in the package) a bit to achieve the data. 

```{r}
data = rbind(
  rddid::sim_data(n = 1000, dgp_type = "cs", seed = 1, jump_t2 = 100) |> filter(time == 2) |> mutate(time = 1),
  rddid::sim_data(n = 1000, dgp_type = "cs", seed = 2, jump_t2 = 200) |> filter(time == 2) |> mutate(time = 2),
  rddid::sim_data(n = 1000, dgp_type = "cs", seed = 3, jump_t2 = 400) |> filter(time == 2) |> mutate(time = 3)
)
data |> arrange(id, time) |> head()
```

To confirm the setup, lets estimate an RD in each time period separately. 

```{r}
cat("RD in period 1: ", round(rdrobust::rdrobust(x = data$R[data$time == 1], y = data$Y[data$time == 1])$coef[1], 1), "\n")
cat("RD in period 2: ", round(rdrobust::rdrobust(x = data$R[data$time == 2], y = data$Y[data$time == 2])$coef[1], 1), "\n")
cat("RD in period 3: ", round(rdrobust::rdrobust(x = data$R[data$time == 3], y = data$Y[data$time == 3])$coef[1], 1), "\n")
```

# Estimate treatment effects

Next, lets estimate the effect in time 3, assuming potential outcome discontinuities evolve linearly over time. 

```{r}
rddid::mult_period_est(df = data, t_star = 2, t_vec = 1, w_vec = 1, h = 600, b = 1200)
```

The output is for conventional (C) and bias-corrected (BC) estimators. The standard errors under all sampling schemes (CS, PC and PV) are outputed--the researcher needs to use the one that fits her application.

