---
title: "How rddid() computes the estimate"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{How rddid() computes the estimate}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 2.8)
```

This article is for readers who want to see the pieces: referees, or users
checking the package against their own code. It rebuilds the RD-DID estimate
and two of the validation tests by hand from the two building blocks,
`rd_bw_cct()` (one period's CCT bandwidths) and `rd_period()` (one period's
local-linear RD), and checks that the package returns the same numbers. It
then runs the four tests on simulated designs in which one assumption fails at
a time. All data are simulated inline, with the truth known. For everyday use,
start with [Get started](rddid-estimation.html).

```{r setup, message = FALSE}
library(rddid)
library(ggplot2)
fmt <- function(x, d = 2) formatC(x, format = "f", digits = d)
```

# Part 1: the estimate

## A simulated design

A treatment of interest `W` and a confounding policy `V` switch at the same
cutoff of a running variable `R`. In the comparison periods `W` is zero for
everybody, so the jump in the outcome at the cutoff in period $t$, $D_t$, is the
confounding jump; in the RD period it is the confounding jump plus the ATT. With
comparison-period weights $w_t$, the estimate is
$$
\widehat{\mathrm{ATT}}(t_{\mathrm{RD}}) = \hat D_{t_{\mathrm{RD}}} - \sum_t w_t \hat D_t ,
$$
where $t_{\mathrm{RD}}$ is the RD period and the sum runs over the comparison
periods. Below, periods 1 and 2 are comparison periods, period 3 is the RD
period, and the confounding jump is 1 in every period.

```{r dgp}
m_a <- function(r, theta) r + (theta / 2) * r^2 * (r >= 0)
dgp_a <- function(n = 2000, conf = c(1, 1, 1), theta = c(2, 2, 2), tau = 1, seed = 1) {
  set.seed(seed)
  R <- runif(n, -1, 1)                  # running variable fixed over time, cutoff at 0
  u <- rnorm(n, 0, 0.5)                 # unit effect
  do.call(rbind, lapply(1:3, function(t) {
    V <- as.integer(R >= 0)             # confounding policy: switches at the cutoff every period
    W <- V * (t == 3)                   # treatment of interest: switches in the RD period only
    data.frame(id = seq_len(n), t = t, R = R,
               Y = m_a(R, theta[t]) + conf[t] * V + tau * W + u + rnorm(n, 0, 0.5))
  }))
}

conf_true <- c(1, 1, 1)   # confounding jump by period
tau_true  <- 1            # ATT in the RD period
dat <- dgp_a(conf = conf_true, tau = tau_true)
head(dat)
```

## Per-period jumps

Each $\hat D_t$ is a standard local-linear RD, estimated one period at a time
with `rd_bw_cct()` (the period's CCT bandwidth pair: main bandwidth $h_t$ and
pilot bandwidth $b_t$) and `rd_period()` (the local-linear fit with the bias
correction and robust variance of Calonico, Cattaneo and Titiunik, 2014). At a
given bandwidth pair, `rd_period()` reproduces the conventional and
bias-corrected estimates and the conventional and robust standard errors of
`rdrobust` (with `vce = "hc1"`) to machine precision.

```{r per-period}
periods <- 1:3
per_period <- lapply(periods, function(t) {
  d  <- dat[dat$t == t, ]
  bw <- rd_bw_cct(d$Y, d$R, c = 0)
  fit <- rd_period(d$Y, d$R, h = unname(bw["h"]), b = unname(bw["b"]), id = d$id, c = 0)
  list(d = d, fit = fit)
})
names(per_period) <- periods

truth <- conf_true + tau_true * (periods == 3)
tab <- data.frame(
  t     = periods,
  D     = sapply(per_period, function(p) p$fit$D),
  SE    = sapply(per_period, function(p) sqrt(p$fit$V_D)),
  D_bc  = sapply(per_period, function(p) p$fit$D_bc),
  SE_bc = sapply(per_period, function(p) sqrt(p$fit$V_D_bc)),
  h     = sapply(per_period, function(p) unname(p$fit$h)),
  n     = sapply(per_period, function(p) p$fit$n),
  truth = truth
)
knitr::kable(tab, digits = 3,
             col.names = c("t", "D", "SE", "D (bc)", "SE (bc)", "h", "n", "truth"))
```

```{r rd-plot, fig.alt = "Binned scatter of Y on R with local-linear fits, one facet per period"}
bin_w <- 0.05
binned <- do.call(rbind, lapply(periods, function(t) {
  d   <- per_period[[as.character(t)]]$d
  bin <- (floor(d$R / bin_w) + 0.5) * bin_w
  agg <- aggregate(d$Y, by = list(R = bin), FUN = mean)
  names(agg)[2] <- "Y"
  agg$t <- t
  agg
}))

fit_lines <- do.call(rbind, lapply(periods, function(t) {
  f <- per_period[[as.character(t)]]$fit
  h <- unname(f$h)
  rbind(
    data.frame(t = t, side = "-", R = c(-h, 0),
               Y = f$sides[["-"]]$beta0 + f$sides[["-"]]$slope * (c(-h, 0) - 0)),
    data.frame(t = t, side = "+", R = c(0, h),
               Y = f$sides[["+"]]$beta0 + f$sides[["+"]]$slope * (c(0, h) - 0))
  )
}))

ggplot(binned, aes(R, Y)) +
  geom_point(size = 0.9, colour = "grey45") +
  geom_line(data = fit_lines, aes(group = side), colour = "firebrick", linewidth = 1.1) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  facet_wrap(~ t, nrow = 1) +
  theme_bw()
```

The comparison-period jumps, $\hat D_1 = `r fmt(tab$D[1])`$ and
$\hat D_2 = `r fmt(tab$D[2])`$, estimate the confounding jump; the RD-period
one, $\hat D_3 = `r fmt(tab$D[3])`$, estimates the confounding jump plus the ATT.

## The aggregation

Averaging the two comparison-period jumps,
$(\hat D_1 + \hat D_2)/2 = `r fmt((tab$D[1] + tab$D[2]) / 2)`$, and subtracting
the average from $\hat D_3 = `r fmt(tab$D[3])`$ gives
`r fmt(tab$D[3] - (tab$D[1] + tab$D[2]) / 2)`. `rddid()` does the same
aggregation. With `bwselect = "cct"` it fits each period at its own
`rd_bw_cct()` bandwidths, as the loop above did (the default rule,
`bwselect = "joint"`, would use one common bandwidth instead):

```{r rddid-call}
res <- rddid(dat, y = "Y", x = "R", time = "t", id = "id", t_rd = 3, bwselect = "cct")
res
```

```{r check-aggregation, include = FALSE}
se_indep <- sqrt(tab$SE[3]^2 + (tab$SE[1]^2 + tab$SE[2]^2) / 4)
stopifnot(
  isTRUE(all.equal(res$estimates["Conventional", "est"], tab$D[3] - mean(tab$D[1:2]))),
  isTRUE(all.equal(res$estimates["Robust", "est"], tab$D_bc[3] - mean(tab$D_bc[1:2]))),
  isTRUE(all.equal(res$estimates["Conventional", "se_cs"], se_indep))
)
```

The `Conventional` row is $\hat D_3 - (\hat D_1 + \hat D_2)/2$ computed from the
`D` column of the table above, and the `Robust (bias-corrected)` row is the same
difference of the `D (bc)` column; both agree with the hand computation to
machine precision. A hand computation of the standard error that treats the
three periods as independent,
$\sqrt{SE_3^2 + (SE_1^2 + SE_2^2)/4} = `r fmt(se_indep, 3)`$ from the `SE`
column, reproduces the repeated cross-section standard error
(`res$estimates["Conventional", "se_cs"]`). The printed standard error,
`r fmt(res$estimates["Conventional", "se"], 3)`, is smaller: the same units enter
every period (the scheme is "panel, running variable fixed over time"), and
`rddid()` adds the cross-period covariances of the jumps. `summary(res)` prints
the robust standard error under all three sampling schemes.

## A confounding jump that moves linearly

```{r linear-conf}
conf_lin <- 0.5 + 0.5 * (1:3)
dat_lin <- dgp_a(conf = conf_lin, tau = tau_true)
res_const <- rddid(dat_lin, y = "Y", x = "R", time = "t", id = "id", t_rd = 3,
                   trend = "constant", bwselect = "cct")
res_const$estimates[, c("est", "se")]
bias_truth <- conf_lin[3] - mean(conf_lin[1:2])
```

Here the confounding jump grows linearly over the periods
(`r paste(conf_lin, collapse = ", ")`). Equal weights net out only the average
of the comparison-period confounding jumps, leaving a bias equal to the year-3
confounding jump minus that average,
`r conf_lin[3]` - (`r conf_lin[1]` + `r conf_lin[2]`)/2 = `r fmt(bias_truth)`,
against an effect of `r tau_true`: the conventional estimate,
`r fmt(res_const$estimates["Conventional", "est"])`, sits near
`r fmt(tau_true + bias_truth)`.

With `trend = "linear"`, `rddid()` extrapolates the line through the
comparison-period jumps to the RD period:

```{r linear-weights}
res_lin <- rddid(dat_lin, y = "Y", x = "R", time = "t", id = "id", t_rd = 3,
                 trend = "linear", bwselect = "cct")
res_lin$weights
res_lin$estimates[, c("est", "se")]
```

```{r linear-arith, include = FALSE}
D1_lin <- res_lin$fits[["1"]]$D
D2_lin <- res_lin$fits[["2"]]$D
D3_lin <- res_lin$fits[["3"]]$D
pred_lin <- sum(res_lin$weights * c(D1_lin, D2_lin))
```

The weights are `r fmt(res_lin$weights[["1"]], 0)` and
`r fmt(res_lin$weights[["2"]], 0)`: the line through
$(1, `r fmt(D1_lin)`)$ and $(2, `r fmt(D2_lin)`)$ reaches `r fmt(pred_lin)` at
$t = 3$, which subtracted from $\hat D_3 = `r fmt(D3_lin)`$ leaves
`r fmt(res_lin$estimates["Conventional", "est"])` (standard error
`r fmt(res_lin$estimates["Conventional", "se"])`), against an effect of
`r tau_true`.

A numeric vector supplies the weights directly, one per comparison period.
Under a constant confounding jump any weights that sum to one are valid; on the
first data set:

```{r custom-weights}
res_custom <- rddid(dat, y = "Y", x = "R", time = "t", id = "id", t_rd = 3,
                    trend = c(0.3, 0.7), bwselect = "cct")
res_custom$estimates[, c("est", "se")]
```

```{r checks-estimate, include = FALSE}
stopifnot(
  abs(res$estimates["Conventional", "est"] - tau_true) < 4 * res$estimates["Conventional", "se"],
  abs(res$estimates["Robust", "est"] - tau_true) < 4 * res$estimates["Robust", "se"],
  abs(res_lin$estimates["Conventional", "est"] - tau_true) < 4 * res_lin$estimates["Conventional", "se"],
  abs(res_lin$estimates["Robust", "est"] - tau_true) < 4 * res_lin$estimates["Robust", "se"]
)
for (tt in periods) {
  stopifnot(abs(tab$D[tab$t == tt] - tab$truth[tab$t == tt]) < 4 * tab$SE[tab$t == tt])
}
```

## Comparison periods where everybody is treated: the ATU

When the treatment of interest `W` is one for everybody in the comparison
periods, rather than zero, the same difference of jumps identifies the ATU, the
effect for the units just below the cutoff. In the paper's treatment of this
case, the ATU design is the ATT design with the two sides of the cutoff
exchanged. The estimate, its standard errors and its bandwidths are the same as
for an ATT call, so `estimand = "atu"` only labels the output; among the four
validation tests, only `rd_compstable()` computes differently (see the ATU
section in Part 2). To show that the estimate is the ATU and not the effect
above the cutoff, the effect of `W` below the cutoff (where `V = 0`) differs here
from its effect above the cutoff (where `V = 1`):

```{r dgp-atu}
dgp_atu <- function(n = 2000, conf = c(1, 1, 1), theta = c(2, 2, 2),
                    tau0 = 1, tau1 = 2, seed = 1) {
  set.seed(seed)
  R <- runif(n, -1, 1)                  # running variable fixed over time, cutoff at 0
  u <- rnorm(n, 0, 0.5)                 # unit effect
  do.call(rbind, lapply(1:3, function(t) {
    V <- as.integer(R >= 0)             # confounding policy: switches at the cutoff every period
    W <- if (t == 3) V else 1           # treatment of interest: everybody treated in comparison periods
    data.frame(id = seq_len(n), t = t, R = R,
               Y = m_a(R, theta[t]) + conf[t] * V +
                 tau0 * W * (1 - V) + tau1 * W * V + u + rnorm(n, 0, 0.5))
  }))
}

tau0 <- 1   # the ATU: effect of W below the cutoff, where V = 0
tau1 <- 2   # effect of W above the cutoff, where V = 1 -- not identified by this design
dat_atu <- dgp_atu(tau0 = tau0, tau1 = tau1)
head(dat_atu)
```

`W` is one for everybody in the comparison periods and equal to `V` in the RD
period, so a comparison-period jump is now the confounding jump among treated
units: the confounding jump of 1 plus `tau1 - tau0`.

```{r rddid-atu-call}
res_atu <- rddid(dat_atu, y = "Y", x = "R", time = "t", id = "id", t_rd = 3,
                 estimand = "atu")
res_atu
tau0
```

The estimate is `r fmt(res_atu$estimates["Conventional", "est"])` (standard
error `r fmt(res_atu$estimates["Conventional", "se"])`), against the true ATU
`tau0` = `r tau0`.

```{r checks-atu, include = FALSE}
est_atu <- res_atu$estimates["Conventional", "est"]
se_atu  <- res_atu$estimates["Conventional", "se"]
stopifnot(
  abs(est_atu - tau0) < 3 * se_atu,
  grepl("ATU", paste(capture.output(res_atu), collapse = "\n"))
)
```

# Part 2: the tests

```{r setup-tests, include = FALSE}
knitr::opts_chunk$set(fig.height = 3)
fmt <- function(x, d = 3) formatC(x, format = "f", digits = d)
fmtp <- function(p) ifelse(p < 0.001, "< 0.001", fmt(p))          # p-values in tables
pmath <- function(p) ifelse(p < 0.001, "p < 0.001", paste("p =", fmt(p)))  # p-values in prose
bin_x <- function(x, w = 0.1) ifelse(x >= 0, floor(x / w) * w + w / 2, -(floor(-x / w) * w + w / 2))
```

## A simulated design with a moving running variable

The four tests apply when the running variable $R$ moves over time within a
unit (a panel whose running variable varies over time); when it is fixed, every
unit is on the same side of the cutoff in every period and the objects below
are degenerate. When $R$ moves, a unit can sit on different sides of the cutoff
in different periods. Its **type** in period $t$ is its side of the cutoff in
the *other* period(s). Write $\pi_{t,(+)}(v)$ and $\pi_{t,(-)}(v)$ for the shares
of type $v$ just above and just below the cutoff in period $t$. The paper's
treatment of a time-varying running variable states four assumptions on these
shares and on the confounding jump within each type, and a test for each. Every test function takes the same long data
frame (`id`, `t`, `R`, `Y`) that `rddid()` takes.

```{r dgp-b}
m_b <- function(r) r + r^2 * (r >= 0)
dgp_b <- function(n = 4000, d = 0, conf = c(1, 1), kappa = 0, sort_p = 0, sort_w = 0.5,
                  tau = 1, periods = 2, gamma = 0, seed = 1) {
  set.seed(seed)
  eta <- rnorm(n)                                          # latent level of the running variable
  R <- sapply(seq_len(periods), function(t) eta + d * (t - 1) + rnorm(n, 0, 0.5))
  if (sort_p > 0) {                                        # sorting: some units just below the RD-period
    V1 <- R[, 1] >= 0; r2 <- R[, periods]                  # cutoff that were above in period 1 move above
    mover <- r2 > -sort_w & r2 < 0 & V1 & runif(n) < sort_p
    R[mover, periods] <- -R[mover, periods]
  }
  V <- R >= 0
  do.call(rbind, lapply(seq_len(periods), function(t) {
    other <- if (periods == 2) V[, 3 - t] else V[, periods]  # the unit's type: side in the other period
    a <- conf[other + 1] + gamma * t                         # confounding jump within the type
    data.frame(id = seq_len(n), t = t, R = R[, t],
               Y = m_b(R[, t]) + a * V[, t] + tau * V[, t] * (t == periods) + kappa * eta + rnorm(n, 0, 0.5))
  }))
}
```

- `d`: a common drift of $R$ between periods; breaks composition stability.
- `conf`: the confounding jump by type, `c(below in the other period, above in
  the other period)`; unequal values break homogeneous confounding.
- `kappa`, `sort_p`: sorting of period-1-above units across the period-2 cutoff,
  with `kappa` making the outcome level depend on the type; breaks type
  continuity.
- `gamma`: a trend in the confounding jump across periods (three-period
  version); breaks constant within-type confounding.

```{r scenarios}
S0   <- dgp_b()                                # every assumption holds
S1   <- dgp_b(d = 0.5)                         # composition changes; confounding homogeneous
S2   <- dgp_b(d = 0.5, conf = c(0.5, 1.5))     # composition changes; confounding heterogeneous
S3   <- dgp_b(kappa = 2, sort_p = 0.5)         # sorting at the RD-period cutoff
S0_3 <- dgp_b(periods = 3)                     # three periods, everything holds
S4   <- dgp_b(periods = 3, gamma = 0.3)        # confounding jump trends across periods
```

```{r type-pattern-plot, fig.alt = "Scatter of R in period 1 vs period 2 for S1, colored by side pattern"}
w1 <- reshape(S1[, c("id", "t", "R")], idvar = "id", timevar = "t", direction = "wide")
names(w1) <- c("id", "R1", "R2")
w1$pattern <- with(w1, paste0(ifelse(R1 >= 0, "+", "-"), ifelse(R2 >= 0, "+", "-")))
ggplot(w1, aes(R1, R2, colour = pattern)) +
  geom_point(size = 0.5, alpha = 0.4) +
  geom_hline(yintercept = 0, linetype = "dashed") + geom_vline(xintercept = 0, linetype = "dashed") +
  labs(colour = "sides in periods 1, 2") + theme_bw()
```

With drift (`d = 0.5`), units just above the period-2 cutoff were mostly below
the cutoff in period 1.

## Type continuity by hand

Type continuity requires $\pi_{t,(+)}(v) = \pi_{t,(-)}(v)$ in every period: the
share of each type jumps by zero at the cutoff. The test regresses, in each
period $t$, the indicator of the unit's side in the other period $s$,
$1\{R_{i,s} \ge 0\}$, on $R_{i,t}$: the local-linear RD jump of that indicator
is $\hat\pi_{t,(+)}(1) - \hat\pi_{t,(-)}(1)$. The two per-period jumps, built
directly with `rd_bw_cct()` and `rd_period()`:

```{r typecont-jumps}
side_share <- function(dat) {
  w <- reshape(dat[, c("id", "t", "R")], idvar = "id", timevar = "t", direction = "wide")
  names(w) <- c("id", "R1", "R2")
  V <- w[, c("R1", "R2")] >= 0
  lapply(1:2, function(t) {
    other <- V[, 3 - t]
    d <- data.frame(id = w$id, R = w[[paste0("R", t)]], ind = as.integer(other))
    bw <- rd_bw_cct(d$ind, d$R, c = 0)
    fit <- rd_period(d$ind, d$R, h = unname(bw["h"]), b = unname(bw["b"]), id = d$id, c = 0)
    list(d = d, fit = fit)
  })
}
ss0 <- side_share(S0); ss3 <- side_share(S3)
tab_ss <- data.frame(
  scenario = rep(c("S0", "S3"), each = 2), t = rep(1:2, 2),
  jump = c(sapply(ss0, function(s) s$fit$D), sapply(ss3, function(s) s$fit$D)),
  se   = c(sapply(ss0, function(s) sqrt(s$fit$V_D)), sapply(ss3, function(s) sqrt(s$fit$V_D)))
)
knitr::kable(tab_ss, digits = 3, col.names = c("scenario", "t", "jump", "SE"))
```

`rd_typecont()` computes the same jumps and combines them into a joint Wald
statistic across periods, with the cross-period covariance of the two jumps
(the same units enter both regressions) inside it. The statistic, its degrees of
freedom and p-value are `$statistic`, `$df`, `$p_value`; the per-period
versions sit in `$per_period[[t]]$ll_wald`:

```{r typecont-wald}
tc0 <- rd_typecont(S0, x = "R", time = "t", id = "id", bwselect = "cct", bc = FALSE)
tc3 <- rd_typecont(S3, x = "R", time = "t", id = "id", bwselect = "cct", bc = FALSE)
tc0_bc <- rd_typecont(S0, x = "R", time = "t", id = "id", bwselect = "cct", bc = TRUE)
```

```{r check-typecont, include = FALSE}
wald_by_hand <- function(ss) sapply(ss, function(s) s$fit$D^2 / s$fit$V_D)
wald_pkg <- function(tc) sapply(c("1", "2"), function(t) tc$per_period[[t]]$ll_wald$stat)
stopifnot(
  isTRUE(all.equal(unname(wald_by_hand(ss0)), unname(wald_pkg(tc0)))),
  isTRUE(all.equal(unname(wald_by_hand(ss3)), unname(wald_pkg(tc3))))
)
```

Each per-period statistic equals the squared jump over its variance from the
table above, to machine precision. For S0,
$\chi^2(`r tc0$df`) = `r fmt(tc0$statistic)`$, $`r pmath(tc0$p_value)`$ (per
period $`r pmath(tc0$per_period[["1"]]$ll_wald$p)`$ and
$`r pmath(tc0$per_period[["2"]]$ll_wald$p)`$); with `bc = TRUE` the Wald uses
the bias-corrected jumps and their robust variance, $`r pmath(tc0_bc$p_value)`$.
For S3, $\chi^2(`r tc3$df`) = `r fmt(tc3$statistic)`$, $`r pmath(tc3$p_value)`$
(per period $`r pmath(tc3$per_period[["1"]]$ll_wald$p)`$ and
$`r pmath(tc3$per_period[["2"]]$ll_wald$p)`$). S0's per-period jumps are
indistinguishable from zero in either period; S3's are not, in either period.

```{r typecont-plot, fig.height = 3.5, fig.alt = "Type-share RD plots, rows S0/S3, columns period 1/2"}
type_share_df <- function(ss, scenario) {
  do.call(rbind, lapply(1:2, function(t) {
    fit <- ss[[t]]$fit; d <- ss[[t]]$d
    agg <- aggregate(d$ind, by = list(R = bin_x(d$R)), FUN = mean)
    names(agg)[2] <- "ind"; agg$scenario <- scenario; agg$t <- t
    h <- unname(fit$h)
    lines <- rbind(
      data.frame(R = c(-h, 0), ind = fit$sides[["-"]]$beta0 + fit$sides[["-"]]$slope * (c(-h, 0) - 0), side = "-"),
      data.frame(R = c(0, h),  ind = fit$sides[["+"]]$beta0 + fit$sides[["+"]]$slope * (c(0, h) - 0),  side = "+")
    )
    lines$scenario <- scenario; lines$t <- t
    list(binned = agg, lines = lines)
  }))
}
b0 <- type_share_df(ss0, "S0"); b3 <- type_share_df(ss3, "S3")
binned <- rbind(do.call(rbind, b0[, "binned"]), do.call(rbind, b3[, "binned"]))
lines  <- rbind(do.call(rbind, b0[, "lines"]),  do.call(rbind, b3[, "lines"]))
ggplot(binned, aes(R, ind)) +
  geom_point(size = 0.9, colour = "grey45") +
  geom_line(data = lines, aes(group = side), colour = "firebrick", linewidth = 1.1) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  facet_grid(scenario ~ t, labeller = label_both) + labs(y = "above the cutoff in the other period") + theme_bw()
```

## Composition stability by hand

Composition stability requires
$\pi_{t_{\mathrm{RD}},(+)}(v) = \pi_{t_0,(+)}(v)$, with $t_0$ a comparison
period: the share of each type among the units just above the cutoff is the same in
the RD period and in the comparison period. The test reflects the
comparison-period above-cutoff units onto a negative axis and stacks them with
the RD-period above-cutoff units. For S1 (RD period 2, comparison period 1):

```{r compstable-build}
w1 <- reshape(S1[, c("id", "t", "R")], idvar = "id", timevar = "t", direction = "wide")
names(w1) <- c("id", "R1", "R2")
above1 <- w1$R1 >= 0; above2 <- w1$R2 >= 0
refl <- rbind(
  data.frame(id = w1$id[above1], x = -(w1$R1[above1]), ind = as.integer(above2[above1])),
  data.frame(id = w1$id[above2], x =   w1$R2[above2],  ind = as.integer(above1[above2]))
)
bw_r <- rd_bw_cct(refl$ind, refl$x, c = 0)
fit_r <- rd_period(refl$ind, refl$x, h = unname(bw_r["h"]), b = unname(bw_r["b"]), id = refl$id, c = 0)
```

```{r compstable-plot, fig.alt = "Binned means and local-linear lines of the reflected composition-stability RD, S1"}
agg_r <- aggregate(refl$ind, by = list(x = bin_x(refl$x)), FUN = mean)
names(agg_r)[2] <- "ind"
h_r <- unname(fit_r$h)
lines_r <- rbind(
  data.frame(x = c(-h_r, 0), ind = fit_r$sides[["-"]]$beta0 + fit_r$sides[["-"]]$slope * (c(-h_r, 0) - 0), side = "-"),
  data.frame(x = c(0, h_r),  ind = fit_r$sides[["+"]]$beta0 + fit_r$sides[["+"]]$slope * (c(0, h_r) - 0),  side = "+")
)
ggplot(agg_r, aes(x, ind)) +
  geom_point(size = 0.9, colour = "grey45") + geom_line(data = lines_r, aes(group = side), colour = "firebrick", linewidth = 1.1) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  labs(x = "reflected running variable", y = "above the cutoff in the other period") + theme_bw()
```

The jump at $x = 0$ estimates $\pi_{2,(+)}(1) - \pi_{1,(+)}(1)$.
`rd_compstable()` runs the same construction and reports the jump, its standard
error and the Wald test:

```{r compstable-wald}
cs0 <- rd_compstable(S0, x = "R", time = "t", id = "id", t_rd = 2, comparisons = 1, bwselect = "cct", bc = FALSE)
cs1 <- rd_compstable(S1, x = "R", time = "t", id = "id", t_rd = 2, comparisons = 1, bwselect = "cct", bc = FALSE)
p1 <- cs1$pairs[["2::1"]]; p0 <- cs0$pairs[["2::1"]]
```

```{r check-compstable, include = FALSE}
stopifnot(isTRUE(all.equal(unname(p1$jumps), fit_r$D)))
```

For S1: jump $= `r fmt(p1$jumps)`$, equal to the hand-built
$`r fmt(fit_r$D)`$; standard error $`r fmt(p1$jump_se)`$;
$\chi^2(`r p1$ll_wald$df`) = `r fmt(p1$ll_wald$stat)`$, $`r pmath(p1$ll_wald$p)`$,
with `r p1$n_trd` above-cutoff units from the RD period, `r p1$n_t0` from the
comparison period, and `r p1$n_both` above the cutoff in both. Those
`r p1$n_both` units appear on both sides of the reflected cutoff. The hand-built
`rd_period()` fit treats the two sides as independent (standard error
$`r fmt(sqrt(fit_r$V_D))`$); the package standard error adds the id-matched
covariance between the two sides. For S0, for contrast: jump
$= `r fmt(p0$jumps)`$, $`r pmath(p0$ll_wald$p)`$.

## ATU designs: estimand = "atu"

With comparison periods in which everybody is treated, the ATU requires
composition stability of the below-cutoff shares,
$\pi_{t_{\mathrm{RD}},(-)}(v) = \pi_{t_0,(-)}(v)$.
`rd_compstable(estimand = "atu")` mirrors the running variable and runs the same
construction on the below-cutoff units, with the same type indicator (above the
cutoff in the other period).

```{r compstable-atu}
cs1_atu <- rd_compstable(S1, x = "R", time = "t", id = "id", t_rd = 2, comparisons = 1,
                         bwselect = "cct", bc = FALSE, estimand = "atu")
p1_atu <- cs1_atu$pairs[["2::1"]]
```

For S1, the `"att"` call above gives jump $= `r fmt(p1$jumps)`$,
$\chi^2(`r p1$ll_wald$df`) = `r fmt(p1$ll_wald$stat)`$, $`r pmath(p1$ll_wald$p)`$,
on `r p1$n_trd` above-cutoff units from the RD period and `r p1$n_t0` from the
comparison period; the `"atu"` call gives jump $= `r fmt(p1_atu$jumps)`$,
$\chi^2(`r p1_atu$ll_wald$df`) = `r fmt(p1_atu$ll_wald$stat)`$,
$`r pmath(p1_atu$ll_wald$p)`$, on `r p1_atu$n_trd` below-cutoff units from the
RD period and `r p1_atu$n_t0` from the comparison period, different units from
the `"att"` call. Both reject under S1's drift.

The other three tests are unchanged under `estimand = "atu"`:

```{r typecont-atu}
tc_att <- rd_typecont(S1, x = "R", time = "t", id = "id", bwselect = "cct", bc = FALSE)
tc_atu <- rd_typecont(S1, x = "R", time = "t", id = "id", bwselect = "cct", bc = FALSE,
                      estimand = "atu")
```

`rd_typecont()` on S1 gives the same joint statistic either way,
$\chi^2(`r tc_att$df`) = `r fmt(tc_att$statistic)`$ under `"att"`
and `r fmt(tc_atu$statistic)` under `"atu"`.

```{r checks-atu-section, include = FALSE}
stat_att <- tc_att$ll_wald$stat
stat_atu <- tc_atu$ll_wald$stat
stopifnot(
  isTRUE(all.equal(stat_att, stat_atu)),
  !isTRUE(all.equal(p1$ll_wald$stat, p1_atu$ll_wald$stat))
)
```

## Homogeneous confounding

Homogeneous confounding requires that, in the comparison period, the
confounding jump is the same for every type. The test
estimates the outcome jump within each type (the unit's side of the cutoff in
the RD period, `type_by = "rd_side"`, the default) in the comparison period, and
tests equality:

```{r homog}
h1 <- rd_homog(S1, y = "Y", x = "R", time = "t", id = "id", t_rd = 2, comparisons = 1, bwselect = "cct", bc = FALSE)
h2 <- rd_homog(S2, y = "Y", x = "R", time = "t", id = "id", t_rd = 2, comparisons = 1, bwselect = "cct", bc = FALSE)
tab_h <- rbind(cbind(scenario = "S1", h1$period_type_jumps), cbind(scenario = "S2", h2$period_type_jumps))
knitr::kable(tab_h[, c("scenario", "period", "type", "jump", "se", "n")], digits = 3, row.names = FALSE)
```

S1: $\chi^2(`r h1$df`) = `r fmt(h1$statistic)`$, $`r pmath(h1$p_value)`$. S2:
$\chi^2(`r h2$df`) = `r fmt(h2$statistic)`$, $`r pmath(h2$p_value)`$. In S1 the
two within-type comparison-period jumps,
$`r fmt(h1$period_type_jumps$jump[1])`$ and
$`r fmt(h1$period_type_jumps$jump[2])`$, are close to each other (both near the
confounding jump of 1); in S2 they are
$`r fmt(h2$period_type_jumps$jump[1])`$ and
$`r fmt(h2$period_type_jumps$jump[2])`$, close to the construction values
`conf = c(0.5, 1.5)`.

```{r homog-plot, fig.alt = "Point-range of within-type jumps, facets S1/S2"}
tab_h$scenario <- factor(tab_h$scenario, levels = c("S1", "S2"))
ggplot(tab_h, aes(type, jump)) +
  geom_pointrange(aes(ymin = jump - 1.96 * se, ymax = jump + 1.96 * se)) +
  facet_wrap(~ scenario) + labs(x = "type (side in the RD period)", y = "within-type jump") + theme_bw()
```

## Constant within-type confounding

Constant within-type confounding requires, for each type, the same confounding
jump in the RD period as in the comparison period. It cannot be
tested in the RD period, where the confounding jump is not seen apart from the
effect; with two comparison periods ($t = 1, 2$, RD period $t = 3$) the test
instead checks whether the within-type jumps are constant *across* the two
comparison periods, a pre-trends check in the difference-in-differences sense:

```{r trendcell}
tr0 <- rd_trendcell(S0_3, y = "Y", x = "R", time = "t", id = "id", t_rd = 3, comparisons = c(1, 2), bwselect = "cct", trend = "constant", bc = FALSE)
tr4 <- rd_trendcell(S4, y = "Y", x = "R", time = "t", id = "id", t_rd = 3, comparisons = c(1, 2), bwselect = "cct", trend = "constant", bc = FALSE)
tab_tr <- rbind(cbind(scenario = "S0_3", tr0$cell_period_jumps), cbind(scenario = "S4", tr4$cell_period_jumps))
knitr::kable(tab_tr[, c("scenario", "cell", "period", "jump", "se", "n")], digits = 3, row.names = FALSE)
```

S0_3: $\chi^2(`r tr0$df`) = `r fmt(tr0$statistic)`$, $`r pmath(tr0$p_value)`$.
S4: $\chi^2(`r tr4$df`) = `r fmt(tr4$statistic)`$, $`r pmath(tr4$p_value)`$. In S4
each type's jump rises from $t = 1$ to $t = 2$: type "+" from
$`r fmt(tr4$cell_period_jumps$jump[1])`$ to
$`r fmt(tr4$cell_period_jumps$jump[2])`$, type "-" from
$`r fmt(tr4$cell_period_jumps$jump[3])`$ to
$`r fmt(tr4$cell_period_jumps$jump[4])`$, by about the construction increment
`gamma = 0.3`. With only two comparison periods the constant form is what can be
tested here; `trend = "linear"` needs three or more.

```{r trendcell-plot, fig.alt = "Point-range of within-type jumps by period, facets S0_3/S4"}
tab_tr$scenario <- factor(tab_tr$scenario, levels = c("S0_3", "S4"))
ggplot(tab_tr, aes(period, jump, colour = cell)) +
  geom_pointrange(aes(ymin = jump - 1.96 * se, ymax = jump + 1.96 * se), position = position_dodge(width = 0.3)) +
  facet_wrap(~ scenario) + labs(colour = "type") + theme_bw()
```

## What each test detects

```{r summary-tests}
tc1 <- rd_typecont(S1, x = "R", time = "t", id = "id", bwselect = "cct", bc = FALSE)
tc2 <- rd_typecont(S2, x = "R", time = "t", id = "id", bwselect = "cct", bc = FALSE)
cs2 <- rd_compstable(S2, x = "R", time = "t", id = "id", t_rd = 2, comparisons = 1, bwselect = "cct", bc = FALSE)
cs3 <- rd_compstable(S3, x = "R", time = "t", id = "id", t_rd = 2, comparisons = 1, bwselect = "cct", bc = FALSE)
h0 <- rd_homog(S0, y = "Y", x = "R", time = "t", id = "id", t_rd = 2, comparisons = 1, bwselect = "cct", bc = FALSE)
h3 <- rd_homog(S3, y = "Y", x = "R", time = "t", id = "id", t_rd = 2, comparisons = 1, bwselect = "cct", bc = FALSE)

rid <- function(dat, ...) rddid(dat, y = "Y", x = "R", time = "t", id = "id", t_rd = max(dat$t), bwselect = "cct", ...)
estfmt <- function(r) sprintf("%s (%s)", fmt(r$estimates["Conventional", "est"]), fmt(r$estimates["Conventional", "se"]))
r0 <- rid(S0); r1 <- rid(S1); r2 <- rid(S2); r3 <- rid(S3)
r0_3 <- rid(S0_3); r4c <- rid(S4); r0_3l <- rid(S0_3, trend = "linear"); r4l <- rid(S4, trend = "linear")

tab1 <- data.frame(
  scenario = c("S0", "S1", "S2", "S3"),
  type_cont_p    = fmtp(c(tc0$p_value, tc1$p_value, tc2$p_value, tc3$p_value)),
  comp_stable_p  = fmtp(c(p0$ll_wald$p, p1$ll_wald$p, cs2$pairs[["2::1"]]$ll_wald$p, cs3$pairs[["2::1"]]$ll_wald$p)),
  homog_p        = fmtp(c(h0$p_value, h1$p_value, h2$p_value, h3$p_value)),
  rddid_est_se   = c(estfmt(r0), estfmt(r1), estfmt(r2), estfmt(r3)),
  truth = 1
)
knitr::kable(tab1, col.names = c("scenario", "type continuity p", "composition stability p",
                                 "homogeneous confounding p", "rddid() est (SE)", "truth"))

tab2 <- data.frame(
  scenario = c("S0_3", "S4"), trendcell_p = fmtp(c(tr0$p_value, tr4$p_value)),
  est_constant = c(estfmt(r0_3), estfmt(r4c)), est_linear = c(estfmt(r0_3l), estfmt(r4l)), truth = 1
)
knitr::kable(tab2, col.names = c("scenario", "constant within-type confounding p",
                                 "rddid() est (SE), constant", "rddid() est (SE), linear", "truth"))
```

```{r zdist, include = FALSE}
zd <- function(r) fmt(abs(r$estimates["Conventional", "est"] - 1) / r$estimates["Conventional", "se"], 1)
```

Distances from the truth below are in standard errors of this one sample. In S0
all four assumptions hold, so the estimate targets the truth; here it is
`r estfmt(r0)`, `r zd(r0)` standard errors from the truth of 1. In S1
composition stability is rejected ($`r pmath(p1$ll_wald$p)`$) while homogeneous
confounding is not ($`r pmath(h1$p_value)`$), and the estimate, `r estfmt(r1)`,
is `r zd(r1)` standard errors from the truth: identification needs only one of
composition stability and homogeneous confounding. In S2 both are rejected
($`r pmath(cs2$pairs[["2::1"]]$ll_wald$p)`$ and $`r pmath(h2$p_value)`$), and
the estimate, `r estfmt(r2)`, is `r zd(r2)` standard errors from the truth: the
composition term. In S3 type continuity is rejected ($`r pmath(tc3$p_value)`$);
the sorting biases the estimate, `r estfmt(r3)`, but in this one sample it is
`r zd(r3)` standard errors from the truth, about as far as S0's, so the estimate
alone does not reveal the problem and the test does. In the three-period design,
S4's test of constant within-type confounding rejects
($`r pmath(tr4$p_value)`$): the constant-trend estimate, `r estfmt(r4c)`, is
`r zd(r4c)` standard errors from the truth, the linear-trend estimate,
`r estfmt(r4l)`, `r zd(r4l)`.

```{r checks-tests, include = FALSE}
stopifnot(
  cs1$pairs[["2::1"]]$ll_wald$p < 0.01, tc3$ll_wald$p < 0.01,
  tc3$per_period[["1"]]$ll_wald$p < 0.01, tc3$per_period[["2"]]$ll_wald$p < 0.01,
  h2$p_value < 0.01, tr4$p_value < 0.05,
  tc0$ll_wald$p > 0.2, tc1$ll_wald$p > 0.2, p0$ll_wald$p > 0.2,
  h1$p_value > 0.2, h0$p_value > 0.2, cs3$pairs[["2::1"]]$ll_wald$p > 0.2, tr0$p_value > 0.2,
  abs(r0$estimates["Conventional", "est"] - 1) < 4 * r0$estimates["Conventional", "se"],
  abs(r1$estimates["Conventional", "est"] - 1) < 4 * r1$estimates["Conventional", "se"],
  abs(r2$estimates["Conventional", "est"] - 1) > 3 * r2$estimates["Conventional", "se"],
  abs(r4c$estimates["Conventional", "est"] - 1) > 3 * r4c$estimates["Conventional", "se"]
)
```

## References

Leventer, D. and D. Nevo (2024). Correcting Invalid Regression Discontinuity
Designs Using Multiple Time-Period Data. arXiv:2408.05847.
<https://arxiv.org/abs/2408.05847>

Calonico, S., M. D. Cattaneo and R. Titiunik (2014). Robust nonparametric
confidence intervals for regression-discontinuity designs. *Econometrica* 82(6),
2295–2326.
