---
title: "Estimating the ATT with rddid()"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Estimating the ATT with rddid()}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 2.8)
```

```{r setup}
library(rddid)
library(ggplot2)
fmt <- function(x, d = 2) formatC(x, format = "f", digits = d)
```

## Setup

A treatment of interest `W` and a confounding treatment `V` both switch at the
same cutoff of a running variable `R`. In comparison periods `W` is uniformly
zero, so the observed discontinuity in period $t$, $D_t$, equals the
confounding discontinuity $\alpha_{t,0}$ (the paper's notation: the jump in
the untreated outcome at the cutoff); in the RD period,
$D_{t_{\mathrm{RD}}} = \mathrm{ATT}(t_{\mathrm{RD}}) + \alpha_{t_{\mathrm{RD}},0}$.
With comparison-period weights $w_t$,
$$
\widehat{\mathrm{ATT}}(t_{\mathrm{RD}}) = \hat D_{t_{\mathrm{RD}}} - \sum_t w_t \hat D_t .
$$

The DGP below has periods $t = 1, 2$ as comparisons, $t = 3$ as the RD period,
and a constant confounding discontinuity `alpha_true` across periods.

```{r dgp}
m_a <- function(r, theta) r + (theta / 2) * r^2 * (r >= 0)
dgp_a <- function(n = 2000, alpha = c(1, 1, 1), theta = c(2, 2, 2), tau = 1, seed = 1) {
  set.seed(seed)
  R <- runif(n, -1, 1)                  # time-invariant running variable, cutoff at 0
  u <- rnorm(n, 0, 0.5)                 # unit effect
  do.call(rbind, lapply(1:3, function(t) {
    V <- as.integer(R >= 0)             # confounding treatment: sharp RD every period
    W <- V * (t == 3)                    # treatment of interest: sharp RD in the RD period only
    data.frame(id = seq_len(n), t = t, R = R,
               Y = m_a(R, theta[t]) + alpha[t] * V + tau * W + u + rnorm(n, 0, 0.5))
  }))
}

alpha_true <- c(1, 1, 1)   # confounding discontinuity, alpha_{t,0}
tau_true   <- 1            # ATT(t_RD)
dat <- dgp_a(alpha = alpha_true, tau = tau_true)
head(dat)
```

The data are long: one row per unit-period, with columns `id`, `t`, `R`, `Y`.
This is the input format every function below expects.

## Per-period discontinuities

Each $\hat D_t$ is a standard local-linear RD, estimated one period at a time
with `rd_bw_cct()` (the per-period CCT bandwidth pair $h_t$, $b_t$) and
`rd_period()` (the local-linear fit with the bias correction and robust
variance of Calonico, Cattaneo and Titiunik 2014). At a given bandwidth pair,
`rd_period()` reproduces the Conventional and Bias-Corrected estimates and the
Conventional and Robust standard errors of `rdrobust` to machine precision.

```{r per-period}
periods <- 1:3
per_period <- lapply(periods, function(t) {
  d  <- dat[dat$t == t, ]
  bw <- rd_bw_cct(d$Y, d$R, c = 0)
  fit <- rd_period(d$Y, d$R, h = unname(bw["h"]), b = unname(bw["b"]), id = d$id, c = 0)
  list(d = d, fit = fit)
})
names(per_period) <- periods

truth <- alpha_true + tau_true * (periods == 3)
tab <- data.frame(
  t     = periods,
  D     = sapply(per_period, function(p) p$fit$D),
  SE    = sapply(per_period, function(p) sqrt(p$fit$V_D)),
  D_bc  = sapply(per_period, function(p) p$fit$D_bc),
  SE_bc = sapply(per_period, function(p) sqrt(p$fit$V_D_bc)),
  h     = sapply(per_period, function(p) unname(p$fit$h)),
  n     = sapply(per_period, function(p) p$fit$n),
  truth = truth
)
knitr::kable(tab, digits = 3,
             col.names = c("t", "D", "SE", "D (bc)", "SE (bc)", "h", "n", "truth"))
```

```{r rd-plot, fig.alt = "Binned scatter of Y on R with local-linear fits, one facet per period"}
bin_w <- 0.05
binned <- do.call(rbind, lapply(periods, function(t) {
  d   <- per_period[[as.character(t)]]$d
  bin <- (floor(d$R / bin_w) + 0.5) * bin_w
  agg <- aggregate(d$Y, by = list(R = bin), FUN = mean)
  names(agg)[2] <- "Y"
  agg$t <- t
  agg
}))

fit_lines <- do.call(rbind, lapply(periods, function(t) {
  f <- per_period[[as.character(t)]]$fit
  h <- unname(f$h)
  rbind(
    data.frame(t = t, side = "-", R = c(-h, 0),
               Y = f$sides[["-"]]$beta0 + f$sides[["-"]]$slope * (c(-h, 0) - 0)),
    data.frame(t = t, side = "+", R = c(0, h),
               Y = f$sides[["+"]]$beta0 + f$sides[["+"]]$slope * (c(0, h) - 0))
  )
}))

ggplot(binned, aes(R, Y)) +
  geom_point(size = 0.9, colour = "grey45") +
  geom_line(data = fit_lines, aes(group = side), colour = "firebrick", linewidth = 1.1) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  facet_wrap(~ t, nrow = 1) +
  theme_bw()
```

The comparison-period discontinuities, $\hat D_1 = `r fmt(tab$D[1])`$ and
$\hat D_2 = `r fmt(tab$D[2])`$, are the confounding $\alpha_{t,0}$; the RD-period
one, $\hat D_3 = `r fmt(tab$D[3])`$, is confounding plus ATT.

## The RD-DID estimate

Averaging the two comparison-period discontinuities,
$(\hat D_1 + \hat D_2)/2 = `r fmt((tab$D[1] + tab$D[2]) / 2)`$, and subtracting
from $\hat D_3 = `r fmt(tab$D[3])`$ gives
`r fmt(tab$D[3] - (tab$D[1] + tab$D[2]) / 2)`. `rddid()` does the same aggregation:

```{r rddid-call}
res <- rddid(dat, y = "Y", x = "R", time = "t", id = "id", t_rd = 3,
             comparisons = c(1, 2), weights = "constant", bwselect = "cct")
res
```

- The two rows, `Conventional` and `Robust`, are the local-linear discontinuity
  estimator and its bias-corrected counterpart, each aggregated with the same
  weights.
- The printed `Std.Err.` and CI are for the sampling scheme on the `scheme`
  line, `res$scheme` = "`r res$scheme`". The three schemes are `cs`
  (repeated cross-section: different units in each period), `pc` (panel with a
  time-constant running variable) and `pv` (panel with a time-varying running
  variable). With `scheme = "auto"` (the default) the scheme is read off the
  data: no `id` gives `cs`; units observed in more than one period, each on
  the same side of the cutoff in every period, give `pc`; any unit on
  different sides of the cutoff in different periods gives `pv`.
- The last line, `Robust SE by scheme`, lists the standard error under all
  three schemes (`res$estimates["Robust", "se_cs"]`, `"se_pc"`, `"se_pv"`);
  here they read `r fmt(res$estimates["Robust", "se_cs"])`,
  `r fmt(res$estimates["Robust", "se_pc"])`, and
  `r fmt(res$estimates["Robust", "se_pv"])`.
- The `weights` line shows `res$weights_type` ("`r res$weights_type`") and the
  values `res$weights`.
- The `bwselect` line reports `"cct"`: each period is fit at its own
  `rd_bw_cct()` bandwidth, so `res$fits` are exactly the per-period fits built
  above. Other bandwidth rules and the sampling schemes are covered in the
  "Bandwidth rules and sampling schemes" vignette.

The scheme describes how the data were sampled; it changes the standard error,
not the estimate. Setting `scheme` replaces the detection, and the printed SE
and CI then correspond to it:

```{r scheme-explicit}
res_pc <- rddid(dat, y = "Y", x = "R", time = "t", id = "id", t_rd = 3,
                comparisons = c(1, 2), weights = "constant", bwselect = "cct", scheme = "pc")
res_pc$estimates["Conventional", c("est", "se", "ci_l", "ci_u")]
```

## Linear confounding trend

```{r linear-alpha}
alpha_lin <- 0.5 + 0.5 * (1:3)
dat_lin <- dgp_a(alpha = alpha_lin, tau = tau_true)
res_const <- rddid(dat_lin, y = "Y", x = "R", time = "t", id = "id", t_rd = 3,
                    comparisons = c(1, 2), weights = "constant", bwselect = "cct")
res_const$estimates[, c("est", "se")]
bias_truth <- alpha_lin[3] - mean(alpha_lin[1:2])
```

With `alpha_lin` growing linearly across periods, constant weights (the equal
average of $\hat D_1, \hat D_2$) net out only the average confounding level,
leaving a bias of $\alpha_3 - (\alpha_1+\alpha_2)/2 = `r fmt(bias_truth)`$
against $\tau = `r fmt(tau_true, 0)`$: the Conventional estimate,
`r fmt(res_const$estimates["Conventional", "est"])`, sits near
$\tau + `r fmt(bias_truth)` = `r fmt(tau_true + bias_truth)`$.

For a linear confounding trend, set `weights = "linear"`: it extrapolates the
line through the comparison-period $\hat D_t$'s to `t_rd`.

```{r linear-weights}
res_lin <- rddid(dat_lin, y = "Y", x = "R", time = "t", id = "id", t_rd = 3,
                  comparisons = c(1, 2), weights = "linear", bwselect = "cct")
res_lin$weights
res_lin$estimates[, c("est", "se")]
```

```{r linear-arith, include = FALSE}
D1_lin <- res_lin$fits[["1"]]$D
D2_lin <- res_lin$fits[["2"]]$D
D3_lin <- res_lin$fits[["3"]]$D
pred_lin <- sum(res_lin$weights * c(D1_lin, D2_lin))
```

`res_lin$weights` is `r fmt(res_lin$weights[["1"]], 0)`, `r fmt(res_lin$weights[["2"]], 0)`:
extrapolating the line through $(1, `r fmt(D1_lin)`)$ and $(2, `r fmt(D2_lin)`)$
to $t = 3$ gives `r fmt(pred_lin)`, subtracted from
$\hat D_3 = `r fmt(D3_lin)`$ to leave
`r fmt(res_lin$estimates["Conventional", "est"])` (SE
`r fmt(res_lin$estimates["Conventional", "se"])`) next to
$\tau = `r fmt(tau_true, 0)`$.

A numeric vector over `comparisons` supplies custom weights. Under a constant
confounding trend any weights summing to one are admissible, e.g. on the
constant-trend data:

```{r custom-weights}
res_custom <- rddid(dat, y = "Y", x = "R", time = "t", id = "id", t_rd = 3,
                    comparisons = c(1, 2), weights = c(0.3, 0.7), bwselect = "cct")
res_custom$estimates[, c("est", "se")]
```

```{r checks, include = FALSE}
stopifnot(
  abs(res$estimates["Conventional", "est"] - tau_true) < 4 * res$estimates["Conventional", "se"],
  abs(res$estimates["Robust", "est"] - tau_true) < 4 * res$estimates["Robust", "se"],
  abs(res_lin$estimates["Conventional", "est"] - tau_true) < 4 * res_lin$estimates["Conventional", "se"],
  abs(res_lin$estimates["Robust", "est"] - tau_true) < 4 * res_lin$estimates["Robust", "se"]
)
for (tt in periods) {
  stopifnot(abs(tab$D[tab$t == tt] - tab$truth[tab$t == tt]) < 4 * tab$SE[tab$t == tt])
}
```

## Comparison periods that are uniformly treated

When the treatment of interest `W` is uniformly one in the comparison periods,
rather than uniformly zero, the same difference of discontinuities identifies
the ATU: Leventer and Nevo, Section 6, show that the ATU design is the ATT
design with the two sides of the cutoff exchanged. The estimate, its standard
errors and its bandwidths are the same as for an ATT call, so
`estimand = "atu"` only labels the output; among the four validation tests,
only `rd_compstable()` computes differently (see the "ATU designs" section of
the [validation-tests vignette](rddid-validation-tests.html#atu-designs-estimand-atu)).

```{r dgp-atu}
dgp_atu <- function(n = 2000, alpha = c(1, 1, 1), theta = c(2, 2, 2),
                    tau0 = 1, tau1 = 2, seed = 1) {
  set.seed(seed)
  R <- runif(n, -1, 1)                  # time-invariant running variable, cutoff at 0
  u <- rnorm(n, 0, 0.5)                 # unit effect
  do.call(rbind, lapply(1:3, function(t) {
    V <- as.integer(R >= 0)             # confounding treatment: sharp RD every period
    W <- if (t == 3) V else 1           # treatment of interest: uniformly on in comparison periods
    data.frame(id = seq_len(n), t = t, R = R,
               Y = m_a(R, theta[t]) + alpha[t] * V +
                 tau0 * W * (1 - V) + tau1 * W * V + u + rnorm(n, 0, 0.5))
  }))
}

tau0 <- 1   # ATU(t_RD): effect of W below the cutoff, where V = 0
tau1 <- 2   # effect of W above the cutoff, where V = 1 -- not identified by this design
dat_atu <- dgp_atu(tau0 = tau0, tau1 = tau1)
head(dat_atu)
```

`tau0` is the ATU, the effect of `W` below the cutoff where `V = 0`; `tau1` is
the effect of `W` above the cutoff where `V = 1`, not identified by this
design. `W` is uniformly one in the comparison periods and equal to `V` in
the RD period, so a comparison period's discontinuity is now
$D_t = \alpha_{t,1} = \alpha_t + \tau_1 - \tau_0$, the confounding
discontinuity among treated units.

```{r rddid-atu-call}
res_atu <- rddid(dat_atu, y = "Y", x = "R", time = "t", id = "id", t_rd = 3,
                  comparisons = c(1, 2), weights = "constant", bwselect = "iter",
                  scheme = "pc", estimand = "atu")
res_atu
tau0
```

The estimate is `r fmt(res_atu$estimates["Conventional", "est"])` (SE
`r fmt(res_atu$estimates["Conventional", "se"])`), against the truth
`tau0` = `r tau0`.

```{r checks-atu, include = FALSE}
est_atu <- res_atu$estimates["Conventional", "est"]
se_atu  <- res_atu$estimates["Conventional", "se"]
stopifnot(
  abs(est_atu - tau0) < 3 * se_atu,
  grepl("ATU", paste(capture.output(res_atu), collapse = "\n"))
)
```
