---
title: "Get started: estimate an RD-DID effect"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get started: estimate an RD-DID effect}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## The setting

A treatment of interest switches on at a cutoff of a running variable in one
period, the **RD period**. A **confounding policy** switches at the same cutoff,
in every period, so the jump in the outcome at the cutoff in the RD period mixes
the treatment effect with the **confounding jump**. In the **comparison
periods** the treatment of interest is uniform at the cutoff (nobody treated, or
everybody treated), so, provided the treatment of interest has no anticipation or carry-over effects there (which the paper assumes), the jump there *is* the confounding jump. `rddid()`
estimates the jump in every period by local-linear RD and subtracts a weighted
average of the comparison-period jumps from the RD-period jump:
$$
\mathrm{ATT}(t_{\mathrm{RD}}) = D_{t_{\mathrm{RD}}} - \sum_t w_t D_t ,
$$
where $t_{\mathrm{RD}}$ is the RD period, $D_t$ is the jump in the outcome at
the cutoff in period $t$, the sum runs over the comparison periods, and $w_t$ is
the weight of comparison period $t$. How the weights are set is the
**confounding-trend assumption**: equal weights if the confounding jump is the
same in every period, a straight-line extrapolation if it moves linearly in time.

## The data

```{r}
library(rddid)
head(rddid_sim)
```

`rddid_sim` is a simulated panel of 1,000 units observed in three years, in
long format: one row per unit and year. `rddid()` uses four columns: the outcome
`Y`, the running variable `R` (cutoff 0), the period `year` and the unit `id`.
The other two record the design: the confounding policy `V` is on above the
cutoff in every year and raises `Y` by 0.5; the treatment of interest `W` is on
above the cutoff in year 3 only and raises `Y` by 1. So year 3 is the RD period,
years 1 and 2 are comparison periods, and **the true effect is 1**.

## Estimate

```{r}
fit <- rddid(rddid_sim, y = "Y", x = "R", time = "year", id = "id", t_rd = 3)
fit
```

```{r, include = FALSE}
fmt <- function(x, d = 2) formatC(x, format = "f", digits = d)
fmtp <- function(p) if (p < 0.001) "below 0.001" else fmt(p, 3)
pp <- summary(fit)$per_period
D <- setNames(pp$jump, pp$period)
```

Line by line:

- **`RD-DID estimate of the ATT in period 3`**: the estimand and the RD period
  (`t_rd = 3`). The ATT is the effect of the treatment on the units just above
  the cutoff, which are the ones treated in the RD period.
- **`Comparison periods: 1, 2 (constant confounding trend; weights 0.5, 0.5)`**:
  by default every period other than `t_rd` is a comparison period. The default
  `trend = "constant"` assumes the confounding jump is the same in every period,
  so each comparison period gets the same weight and their average stands in for
  the confounding jump in year 3.
- **`Sampling scheme`**: how the data were sampled, read off `id` and `R`. Here
  it is a panel (the same units in every year) whose running variable does not
  move. The scheme sets which standard error is reported: in a panel the same
  units enter every year's fit, so the yearly jumps are correlated, and the
  standard error takes that into account. Under the `"joint"` (default) and
  `"iter"` bandwidth rules the scheme also enters the bandwidth (next line),
  which weighs bias against the variance of the estimate, and so it can move the
  estimate; only with a fixed `h` or `bwselect = "cct"` does it leave the
  estimate untouched.
- **`Bandwidth: common h`**: the default rule, `bwselect = "joint"`, uses one
  main bandwidth `h` in every period, chosen to minimize the asymptotic mean
  squared error of the RD-DID estimate. **`Pilot bandwidth b (period = value)`**
  lists the bandwidth of the bias correction in each period, the RD period first.
- **The two rows.** `Conventional` is the local-linear estimate with its
  conventional standard error. `Robust (bias-corrected)` is the bias-corrected
  estimate with its robust standard error (Calonico, Cattaneo and Titiunik,
  2014). Each row has its z statistic, p-value and 95% confidence interval.

The true effect is 1; the two estimates are
`r fmt(coef(fit)[["Conventional"]])` and `r fmt(coef(fit)[["Robust"]])`. A plain
RD in year 3 alone would estimate the effect plus the confounding jump,
1 + 0.5 = 1.5. The year-3 fit is stored in the object:

```{r}
fit$fits[["3"]]
```

Its jump, `r fmt(D[["3"]])`, is that naive RD estimate, whose target is 1.5,
not 1.
`rddid()` subtracts the average of the year-1 and year-2 jumps,
(`r fmt(D[["1"]])` + `r fmt(D[["2"]])`) / 2 = `r fmt((D[["1"]] + D[["2"]]) / 2)`,
its estimate of the confounding jump of 0.5.

## Summary, coefficients and tables

```{r}
summary(fit)
```

`summary()` adds the per-period fits. Each row gives the period's role, its
coefficient in the estimate (+1 for the RD period, minus its weight for a
comparison period), its number of observations, its bandwidths, and its jump
with standard error, conventional (`jump`, `s.e.`) and bias-corrected
(`jump (bc)`, `s.e. (rb)`). The estimate is the sum of coefficient times jump:
`r fmt(D[["3"]], 3)` - 0.5 × `r fmt(D[["1"]], 3)` - 0.5 × `r fmt(D[["2"]], 3)`
= `r fmt(coef(fit)[["Conventional"]], 3)`.

The last line gives the robust standard error under each of the three sampling
schemes. The printout uses the one that matches how the data were sampled; the
others are there for comparison, not to choose from. Here the cross-section
standard error, `r fmt(fit$estimates["Robust", "se_cs"], 3)`, ignores that the
same units appear in every year; the panel one, which accounts for it, is
`r fmt(fit$estimates["Robust", "se_pc"], 3)`.

```{r}
coef(fit)
confint(fit)
generics::tidy(fit)
```

`coef()` returns the two point estimates and `confint()` their confidence
intervals. `tidy()` (from the generics package) returns one row per estimate, so
table makers such as modelsummary work with `rddid()` fits.

## A confounding jump that moves linearly

```{r}
fit_lin <- rddid(rddid_sim, y = "Y", x = "R", time = "year", id = "id", t_rd = 3,
                 trend = "linear")
fit_lin
```

With `trend = "linear"` the confounding jump may move linearly in time. The line
through the year-1 and year-2 jumps is extrapolated to year 3, which gives
weights -1 and 2: the confounding jump in year 3 is estimated by
$2 D_2 - D_1$. This needs at least two comparison periods. The standard error
grows (`r fmt(fit_lin$estimates["Conventional", "se"], 3)` against
`r fmt(fit$estimates["Conventional", "se"], 3)`) because extrapolating a line
amplifies noise. In `rddid_sim` the confounding jump is constant, so both
estimates are valid; the linear one gives up precision to allow for a linear
trend. A numeric vector instead of `"constant"` or `"linear"` sets the weights
directly, one per comparison period.

## Comparison periods where everybody is treated: the ATU

When everybody is treated in the comparison periods, rather than nobody, the
same difference of jumps estimates the ATU: the effect for the units just below
the cutoff, which are untreated in the RD period. Set `estimand = "atu"`. The
estimate, standard errors and bandwidths are the same as without it, so the
argument labels the output (among the tests it changes only
`rd_compstable()`). For illustration, make everybody treated in years 1 and 2
of `rddid_sim`, with the same effect of 1:

```{r}
sim_atu <- transform(rddid_sim, W = ifelse(year < 3, 1, W), Y = Y + (year < 3))
rddid(sim_atu, y = "Y", x = "R", time = "year", id = "id", t_rd = 3, estimand = "atu")
```

The numbers are those of `fit`, now labeled the ATU; the true ATU here is 1.

## When the running variable moves over time

In `rddid_sim` every unit stays on its side of the cutoff. In many applications
the running variable moves (a population count, a test score, an income), and
some units, the **switchers**, are above the cutoff in one period and below it
in another. The units at the cutoff in the RD period can then be a different mix
from the units at the cutoff in a comparison period, and the difference of jumps
can be biased. A unit's **type** is the side of the cutoff it is on in the other
period(s). The RD-DID estimate identifies the ATT when the type distribution is
continuous at the cutoff and the confounding jump within each type is constant
over time, and in addition either the type composition is stable across periods
or the confounding jump is the same for every type (Leventer and Nevo, 2024).
Four tests check these assumptions. `rddid_sim_pv` is the `rddid_sim` design
with a running variable that drifts between years.

**Type continuity.** H0: the share of each type jumps by zero at the cutoff, in
every period. A rejection means units sort around the cutoff by their side in
other periods, so the jump in the outcome partly reflects who the units are; the
estimate can then be biased, whatever the other tests say.

```{r}
tc <- rd_typecont(rddid_sim_pv, x = "R", time = "year", id = "id")
tc
```

With three years, a unit's type in one year is its pair of sides in the other
two, in time order: `++` is above the cutoff in both, `+-` above in the earlier
and below in the later, and so on.

**Composition stability.** H0: the share of each type among the units just above
the cutoff is the same in the RD period and in each comparison period. A rejection
means the units at the cutoff in the RD period are a different mix of types from
those in a comparison period; the estimate is then valid only if the next test's
null holds.

```{r}
cs <- rd_compstable(rddid_sim_pv, x = "R", time = "year", id = "id", t_rd = 3)
cs
```

**Homogeneous confounding.** H0: in each comparison period the confounding jump
is the same for every type. Here, and in the next test, a unit's type is its
side of the cutoff in the RD period (`+` above, `-` below). A rejection means
the confounding jump differs by type; together with a rejection of composition stability, the comparison-period
jumps do not measure the confounding jump of the units at the RD-period cutoff,
and the estimate is biased.

```{r}
hg <- rd_homog(rddid_sim_pv, y = "Y", x = "R", time = "year", id = "id", t_rd = 3)
hg
```

**Constant within-type confounding.** H0: within each type, the confounding jump
is the same in every comparison period. This is a pre-trends check: the
assumption concerns the RD period, where the confounding jump cannot be seen
apart from the effect, so the test asks whether it is stable across the
comparison periods. A rejection means the constant-trend weights do not cancel
the confounding jump; `trend = "linear"` is the alternative when the movement
looks linear.

```{r}
tr <- rd_trendcell(rddid_sim_pv, y = "Y", x = "R", time = "year", id = "id", t_rd = 3)
tr
```

One table for all four:

```{r}
do.call(rbind, lapply(list(tc, cs, hg, tr), generics::tidy))
```


In `rddid_sim_pv` three of the four nulls are true by construction: types are
continuous at the cutoff, and the confounding jump is 0.5 for every unit in
every year. Composition stability is false: units at the year-3 cutoff were
mostly on the same side in years 1 and 2, while units at the year-1 cutoff are
spread evenly over the four types. In this sample the four p-values are
`r fmtp(tc$p_value)`, `r fmtp(cs$p_value)`, `r fmtp(hg$p_value)` and
`r fmtp(tr$p_value)` (in the order of the table). A p-value below 0.05 for a true null is a false
rejection, which a test at the 5% level is designed to make in about one sample in twenty. The article on
the identification assumptions reads each printout in detail.

## What to report

- **The estimate.** Both rows, as the paper does. Base the confidence interval
  and p-value on the `Robust (bias-corrected)` row: at a bandwidth chosen to
  minimize the mean squared error, the conventional interval is centered on a
  biased estimate and can under-cover (Calonico, Cattaneo and Titiunik, 2014).
- **The design.** The RD period, the comparison periods and the trend assumption
  (with the weights), the bandwidth rule and `h`, and the sampling scheme. All
  are in the first lines of the printout; `generics::glance(fit)` returns them
  as a one-row table.
- **The tests**, when the running variable moves over time: the table above.

## Next

- [How rddid() computes the estimate](rddid-how-it-works.html): the estimate and
  the tests rebuilt by hand, for readers who want to see the pieces.
- [Bandwidth rules and sampling schemes](rddid-options.html): `bwselect`, fixed
  bandwidths, and the standard error under each scheme.
- [Checking the identification assumptions](rddid-validation-tests.html): the
  four tests one at a time.
- [Pictures of the checks](rddid-plots.html): `plot()` on a fit and on each test,
  and the switchers picture.

## References

Leventer, D. and D. Nevo (2024). Correcting Invalid Regression Discontinuity
Designs Using Multiple Time-Period Data. arXiv:2408.05847.
<https://arxiv.org/abs/2408.05847>

Calonico, S., M. D. Cattaneo and R. Titiunik (2014). Robust nonparametric
confidence intervals for regression-discontinuity designs. *Econometrica* 82(6),
2295–2326.
