---
title: "rddid-constant"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{rddid-constant}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Intro

This is an example notebook of how to estimate treatment effects in RDDID designs with a constant potential outcome discontinuities assumption. The identification framework is based on Corollary 1(i) in Leventer and Nevo (2024). The estimation framework is presented in Section 6 in the paper. 

First, lets install the rddid package.

```{r setup}
library(rddid)
```

# Setup

To create an easy example that can be replicated, we use the simulated data in the paper. 

For now, lets go with repeated cross section (CS) sampling scheme.

```{r}
data = rddid::sim_data(n = 1000, dgp_type = "cs", seed = 42)
head(data)
```

# Estimate treatment effects

Next, lets estimate the effect in time 2, assuming potential outcome discontinuitites are constant across time. 

```{r}
rddid::mult_period_est(df = data, t_star = 2, t_vec = 1, w_vec = 1, h = 600, b = 1200)
```

The output is for conventional (C) and bias-corrected (BC) estimators. The standard errors under all sampling schemes (CS, PC and PV) are outputed--the researcher needs to use the one that fits her application.

