---
title: "Checking the identification assumptions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Checking the identification assumptions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
fmt <- function(x, d = 3) formatC(x, format = "f", digits = d)
pv <- function(p) if (p < 0.001) "p < 0.001" else paste("p =", fmt(p))
```

[Get started](rddid-estimation.html) runs the four tests and reads their
printouts. This article adds how each statistic is built, the tests under the
ATU design, their options, and how the four results read together.

## When the tests apply

The tests are for a panel in which the running variable moves over time, so that
some units, the **switchers**, are above the cutoff in one period and below it in
another. The units at the cutoff in the RD period can then be a different mix
from the units at the cutoff in a comparison period, and the difference of the
two jumps can mix the treatment effect with that difference in who is at the
cutoff. A unit's **type** records its side of the cutoff in other periods; the
table in [Get started](rddid-estimation.html#when-the-running-variable-moves-over-time)
gives the two versions the tests use. The RD-DID estimate identifies the ATT when

1. **type continuity** holds,
2. **constant within-type confounding** holds, and
3. **composition stability** or **homogeneous confounding** holds (one of the
   two is enough).

This is the paper's result for a time-varying running variable (Leventer and
Nevo, 2024). There is one test per assumption. When the running variable does
not move, as in `rddid_sim`, every unit has the same type in every period and
the tests are not informative; in a repeated cross-section the types are not
observed.

The examples use `rddid_sim_pv`: the `rddid_sim` design (a confounding jump of
0.5 in every year, a treatment effect of 1 in year 3, the RD period) with a
running variable that drifts between years. Most units stay on one side;
some switch:

```{r}
library(rddid)
above <- rddid_sim_pv$R >= 0
switcher <- tapply(above, rddid_sim_pv$id, function(a) length(unique(a)) > 1)
sum(switcher)
```

Every test takes the same long data frame and column names as `rddid()`. By
default each test uses the bias-corrected jumps with their robust variances
(`bc = TRUE`), like the `Robust (bias-corrected)` row of `rddid()`, and picks a
separate CCT bandwidth for every local-linear fit it runs (`bwselect = "cct"`).

## Type continuity: `rd_typecont()`

**H0: the share of each type jumps by zero at the cutoff, in every period.**

```{r}
tc <- rd_typecont(rddid_sim_pv, x = "R", time = "year", id = "id", t_rd = 3)
tc
```

- `Types: ++, +-, -+, --`: a unit's sides of the cutoff in the other two years,
  in time order (the first row of the type table in Get started).
- The test runs, in each year and for each type, a local-linear RD of the
  indicator "the unit is of this type" on the running variable; the jump is the
  difference between the type's share just above and just below the cutoff.
- `Joint Wald chi-squared(9)`: in each year the four shares sum to one, so three
  jumps are free; three years give 9. The `Period` lines test each year alone.

Here `r pv(tc$p_value)`. Type continuity holds in `rddid_sim_pv` by
construction, so a p-value below 0.05 would be a false rejection.

**If it rejects:** units sort around the cutoff according to their side in other
periods, so the jump in the outcome partly reflects a jump in who the units are.
The RD-DID estimate can then be biased, whatever the other tests say.

## Composition stability: `rd_compstable()`

**H0: the share of each type among the units just above the cutoff is the same in the
RD period and in each comparison period.**

```{r}
cs <- rd_compstable(rddid_sim_pv, x = "R", time = "year", id = "id", t_rd = 3)
cs
```

- One block per pair of the RD period and a comparison period: `Pair 3::1`
  compares year 3 with year 1. The types are those of type continuity (a
  unit's sides in the other two years), so there are four types and three free
  shares (`chi-squared(3)`).
- The test stacks the units above the cutoff in the two years, with the
  comparison year's running variable reflected below an artificial cutoff, and
  tests the jump in each type share there.
- `n above the cutoff`: the units above the cutoff in the RD year, in the
  comparison year, and in both. Units above the cutoff in both years appear on
  both sides of the stacked regression; the standard error accounts for them.
- `Joint over pairs`: the sum of the pair statistics and degrees of freedom. It treats
  the pairs as independent although they share the RD-period units, so its
  p-value is approximate; the paper's test is per pair.

Here `r pv(cs$p_value)`. Composition stability is false in `rddid_sim_pv`
by construction: the units at the year-3 cutoff were mostly on the same side in
years 1 and 2, while the units at the year-1 cutoff are spread evenly over the
four types. A rejection here is a correct detection; with 1,000 units the test
does not always detect it.

**If it rejects:** the units at the cutoff in the RD period are a different mix
of types from those in a comparison period. If the confounding jump also differs
across types, the comparison-period jump measures the confounding jump of a
different mix, and the estimate can be biased. If the confounding jump is the same
for every type (the next test), the mix does not matter and the estimate is
still valid.

## Homogeneous confounding: `rd_homog()`

**H0: in each comparison period the confounding jump is the same for every
type.**

Here a unit's type is its side of the cutoff in the RD period (the second row
of the type table in Get started; `type_by = "rd_side"`, the default).

```{r}
hg <- rd_homog(rddid_sim_pv, y = "Y", x = "R", time = "year", id = "id", t_rd = 3)
hg
```

- The table gives, in each comparison year, the jump in the outcome at the cutoff
  among the units below (`-`) and above (`+`) the cutoff in year 3. In a
  comparison period the jump is the confounding jump, so these are the
  confounding jumps by type.
- `Wald chi-squared(2)`: one contrast (type `+` against the reference type `-`)
  per comparison year.

Here `r pv(hg$p_value)`. The null is true in `rddid_sim_pv` (the
confounding jump is 0.5 for every unit), so a p-value below 0.05 would be a false
rejection. The per-type jumps are imprecise (standard errors
`r fmt(min(hg$period_type_jumps$se), 2)` to
`r fmt(max(hg$period_type_jumps$se), 2)`).

**If it rejects:** the confounding jump differs across types. If composition
stability holds, a confounding jump that differs across types does not bias the
estimate. If both nulls are false, the estimate can be biased.

## Constant within-type confounding: `rd_trendcell()`

**H0: within each type, the confounding jump is the same in every comparison
period.**

```{r}
tr <- rd_trendcell(rddid_sim_pv, y = "Y", x = "R", time = "year", id = "id", t_rd = 3)
tr
```

- The per-cell jumps are those of `rd_homog()`, compared the other way:
  `rd_homog()` compares types within a comparison period, `rd_trendcell()`
  compares comparison periods within a type.
- `Wald chi-squared(2)`: one contrast (year 2 against year 1) per type.

This is a pre-trends check. The assumption concerns the RD period, where the
confounding jump cannot be separated from the effect, so the test asks whether
the within-type confounding jump is stable across the comparison periods. Not
rejecting does not establish the assumption in the RD period. Here
`r pv(tr$p_value)`.

With `trend = "linear"` the null is that the within-type confounding jump moves
linearly across the comparison periods. That needs at least three comparison
periods; with two, the function says so and returns no statistic.

**If it rejects:** the within-type confounding jump changes over time, and
constant weights do not cancel it. `rddid(trend = "linear")` allows a
within-type confounding jump that moves linearly in time, and
`rd_trendcell(trend = "linear")` tests that version (three or more comparison
periods).

## ATU designs

When everybody is treated in the comparison periods (`estimand = "atu"`), the
estimate concerns the units just below the cutoff, and composition stability
concerns the below-cutoff shares. `rd_compstable(estimand = "atu")` mirrors the
running variable and runs the same test on the units below the cutoff; the
other three tests give the same numbers with either estimand, so `estimand` only
labels them. `rddid_sim_pv` is an ATT design, so the call below only shows the
mechanics:

```{r}
rd_compstable(rddid_sim_pv, x = "R", time = "year", id = "id", t_rd = 3, estimand = "atu")
```

In this printout the H0 line and the counts refer to the units below the cutoff:
under the ATU design those are the units untreated in the RD period, whose
composition the estimate relies on.

## Run all four and tabulate

```{r}
tests <- list(
  rd_typecont(rddid_sim_pv, x = "R", time = "year", id = "id", t_rd = 3),
  rd_compstable(rddid_sim_pv, x = "R", time = "year", id = "id", t_rd = 3),
  rd_homog(rddid_sim_pv, y = "Y", x = "R", time = "year", id = "id", t_rd = 3),
  rd_trendcell(rddid_sim_pv, y = "Y", x = "R", time = "year", id = "id", t_rd = 3)
)
do.call(rbind, lapply(tests, tidy))
```

`tidy()` returns one row per test, so the four stack
into one table for a paper or a table maker such as modelsummary. The
composition-stability row is the joint-over-pairs statistic, whose p-value is
approximate.

Together, the four results map onto the list at the top. A rejection of type
continuity, or of constant within-type confounding, would call the estimate into
question; so would a rejection of both composition stability and homogeneous
confounding. A rejection of only one of those two would not on its own, because
either one is enough for identification. In `rddid_sim_pv` the truth is known: composition stability fails and
the other three assumptions hold, so the estimate is valid:

```{r}
rddid(rddid_sim_pv, y = "Y", x = "R", time = "year", id = "id", t_rd = 3)
```

The true effect is 1.

## Options

- `bwselect`: `"cct"` (default) chooses a CCT bandwidth for every local-linear
  fit the test runs; `"rot"` uses a rule of thumb. `h` fixes the bandwidth.
- `bc`: `TRUE` (default) tests the bias-corrected jumps with their robust
  variances; `FALSE` uses the conventional ones.
- `scheme`: detected from the data as in `rddid()`; set it to override.
- `rd_homog()` and `rd_trendcell()`: `type_by = "rd_side"` (default, the side in
  the RD period) or `"pattern"` (the sides in all other periods); `min_n` drops
  cells with too few observations on either side of the cutoff, with a message.

The article [How rddid() computes the estimate](rddid-how-it-works.html)
rebuilds the type-continuity and composition-stability tests by hand and runs
all four on simulated designs in which one assumption fails at a time, with the
resulting bias in the estimate.

## Reference

Leventer, D. and D. Nevo (2024). Correcting Invalid Regression Discontinuity
Designs Using Multiple Time-Period Data. arXiv:2408.05847.
<https://arxiv.org/abs/2408.05847>
