---
title: "Bandwidth rules and sampling schemes"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Bandwidth rules and sampling schemes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 3)
```

## Setup

This vignette covers the `rddid()` arguments that control the bandwidths
($h_t$, the period-$t$ bandwidth, and $b_t$, the pilot bandwidth for the
bias correction) and the standard errors reported under the three sampling
schemes. For the estimator, the weights, and the per-period output, see
`vignette("rddid-estimation")`.

Both parts use the same one-sided-curvature DGP: periods 1 and 2 are
comparison periods, period 3 is the RD period, $\alpha_t = 1$ in every
period and $\tau = 1$, so $D_1 = D_2 = 1$, $D_3 = 2$, $\mathrm{ATT} = 1$.
Part 1 varies the curvature $\theta_t$ across periods so that the
per-period bandwidths differ; Part 2 varies the sampling scheme.

```{r}
library(rddid)
fmt <- function(x, d = 3) formatC(x, format = "f", digits = d)
# per-period bandwidths h_t of a fitted rddid object, ordered by period
# (r$fits lists the RD period first)
bw_by_period <- function(r) {
  h <- vapply(r$fits, function(f) unname(f$h), numeric(1))
  round(h[order(as.numeric(names(h)))], 3)
}
# bandwidth method, h_t by period, and the estimates
peek <- function(r) {
  cat("method:", r$bandwidth$method, "\n"); print(bw_by_period(r))
  print(round(r$estimates[, c("est", "se")], 3))
}

m_a <- function(r, theta) r + (theta / 2) * r^2 * (r >= 0)
dgp_a <- function(n = 2000, alpha = c(1, 1, 1), theta = c(2, 2, 2), tau = 1,
                  scheme = c("pc", "cs", "pv"), seed = 1) {
  scheme <- match.arg(scheme)
  set.seed(seed)
  R0 <- runif(n, -1, 1)                 # running variable at baseline, cutoff at 0
  u0 <- rnorm(n, 0, 0.5)                # unit effect
  do.call(rbind, lapply(1:3, function(t) {
    if (scheme == "cs") {               # new units every period
      R <- runif(n, -1, 1); u <- rnorm(n, 0, 0.5); id <- (t - 1) * n + seq_len(n)
    } else {                            # same units every period
      R <- if (scheme == "pv") R0 + rnorm(n, 0, 0.1) else R0
      u <- u0; id <- seq_len(n)
    }
    V <- as.integer(R >= 0)             # confounding treatment: sharp RD every period
    W <- V * (t == 3)                    # treatment of interest: RD period only
    data.frame(id = id, t = t, R = R,
               Y = m_a(R, theta[t]) + alpha[t] * V + tau * W + u + rnorm(n, 0, 0.5))
  }))
}
```

## Bandwidth rules

Data for this section have a curvature $\theta_t$ that differs across
periods, under the PC sampling scheme:

```{r}
dat1 <- dgp_a(theta = c(1, 6, 3))
```

### `bwselect = "cct"`

Per-period CCT/IK bandwidths: each $h_t$ (and pilot $b_t$) minimizes the
MSE of its own $\widehat D_t$, ignoring the other periods.

```{r}
r_cct <- rddid(dat1, y = "Y", x = "R", time = "t", id = "id", t_rd = 3, bwselect = "cct")
peek(r_cct)
```

The per-period $h_t$ sit in `r_cct$fits[[k]]$h`: $h_1 =
`r fmt(r_cct$fits[["1"]]$h)`$, $h_2 = `r fmt(r_cct$fits[["2"]]$h)`$,
$h_3 = `r fmt(r_cct$fits[["3"]]$h)`$; the narrowest is period 2, the most
curved.

### `bwselect = "joint"`

A single common $h^\ast$ minimizes the asymptotic MSE of the aggregate
$\widehat{\mathrm{ATT}}(t_{\mathrm{RD}})$, using one bandwidth for every
period.

```{r}
r_joint <- rddid(dat1, y = "Y", x = "R", time = "t", id = "id", t_rd = 3, bwselect = "joint")
peek(r_joint)
```

`r_joint$bandwidth$h` is the common $h^\ast = `r fmt(r_joint$bandwidth$h)`$,
applied to every period's `fits[[k]]$h`.

### `bwselect = "iter"` (the default)

Period-specific bandwidths targeting the same aggregate MSE as `"joint"`,
but with one $h_t$ per period, found by coordinate descent started at
$h^\ast$.

```{r}
r_iter <- rddid(dat1, y = "Y", x = "R", time = "t", id = "id", t_rd = 3, bwselect = "iter")
peek(r_iter)
```

`start` sets the seed of the descent: `start = "hstar"` (default) starts
every period at the common $h^\ast$; `start = "cct"` starts each period at
its own CCT/IK $h_t$; or a named vector supplies a manual per-period seed.

```{r}
r_iter_cct <- rddid(dat1, y = "Y", x = "R", time = "t", id = "id", t_rd = 3,
                    bwselect = "iter", start = "cct")
bw_by_period(r_iter_cct)
```

`regularize` (default `TRUE`) adds an `rdrobust`-style regularization term
to the squared-bias part of the `"joint"` and `"iter"` objectives, so a
near-zero estimated curvature does not blow the bandwidth up; `reg_const`
(default `r eval(formals(rddid)$reg_const)`) is its constant. $h$ and $b$
can also be fixed by hand, bypassing bandwidth selection entirely:

```{r}
r_fixed <- rddid(dat1, y = "Y", x = "R", time = "t", id = "id", t_rd = 3, h = 0.3)
r_fixed$bandwidth$method
```

```{r, results = "asis"}
estse <- function(r, row) paste0(fmt(r$estimates[row, "est"]), " (", fmt(r$estimates[row, "se"]), ")")
tab <- data.frame(
  rule = c("cct", "joint", "iter"),
  h_1 = c(r_cct$fits[["1"]]$h, r_joint$bandwidth$h, r_iter$fits[["1"]]$h),
  h_2 = c(r_cct$fits[["2"]]$h, r_joint$bandwidth$h, r_iter$fits[["2"]]$h),
  h_3 = c(r_cct$fits[["3"]]$h, r_joint$bandwidth$h, r_iter$fits[["3"]]$h),
  Conventional = sapply(list(r_cct, r_joint, r_iter), estse, row = "Conventional"),
  Robust = sapply(list(r_cct, r_joint, r_iter), estse, row = "Robust"),
  truth = 1
)
knitr::kable(tab, digits = 3)
```

The three rules give different per-period bandwidths — from
$h_3 = `r fmt(r_cct$fits[["3"]]$h)`$ under `"cct"` down to
$h_3 = `r fmt(r_iter$fits[["3"]]$h)`$ under `"iter"` — and, on these data,
Conventional estimates within one standard error of each other
(`r fmt(r_cct$estimates["Conventional", "est"])` to
`r fmt(r_iter$estimates["Conventional", "est"])`, each with SE around
`r fmt(r_cct$estimates["Conventional", "se"])`).

## Sampling schemes

CS is repeated cross-section (different units each period), PC is a panel
with a time-constant running variable, and PV is a panel with a
time-varying running variable. The three differ in the standard error:
under PC the same units enter every period's local regression, so the
per-period discontinuity estimates are correlated through the unit effect;
under PV units move across the window, so the shared-unit covariance is of
smaller order than the variance; under CS there is no cross-period
covariance.

For the CS data, `id = NULL` treats every row as a distinct unit:

```{r}
dat_cs <- dgp_a(scheme = "cs")
dat_pc <- dgp_a(scheme = "pc")
dat_pv <- dgp_a(scheme = "pv")
r_cs <- rddid(dat_cs, y = "Y", x = "R", time = "t", id = NULL, t_rd = 3, bwselect = "cct")
r_pc <- rddid(dat_pc, y = "Y", x = "R", time = "t", id = "id", t_rd = 3, bwselect = "cct")
r_pv <- rddid(dat_pv, y = "Y", x = "R", time = "t", id = "id", t_rd = 3, bwselect = "cct")
c(cs = r_cs$scheme, pc = r_pc$scheme, pv = r_pv$scheme)
```

```{r, results = "asis"}
se_tab <- data.frame(
  data = c("cs", "pc", "pv"),
  rbind(r_cs$estimates["Conventional", c("se_cs", "se_pc", "se_pv")],
        r_pc$estimates["Conventional", c("se_cs", "se_pc", "se_pv")],
        r_pv$estimates["Conventional", c("se_cs", "se_pc", "se_pv")]))
knitr::kable(se_tab, digits = 3, row.names = FALSE)
```

On the PC data, `se_pc` (`r fmt(r_pc$estimates["Conventional", "se_pc"])`)
is smaller than `se_cs` (`r fmt(r_pc$estimates["Conventional", "se_cs"])`);
on the PV data, `se_pv` (`r fmt(r_pv$estimates["Conventional", "se_pv"])`)
is close to `se_cs` (`r fmt(r_pv$estimates["Conventional", "se_cs"])`)
while `se_pc` (`r fmt(r_pv$estimates["Conventional", "se_pc"])`) is
smaller; on the CS data all three coincide, at
`r fmt(r_cs$estimates["Conventional", "se_cs"])`.

Passing `scheme` explicitly picks which of the three the print method's SE
and CI correspond to, rather than detecting it from the data:

```{r}
r_pv_explicit <- rddid(dat_pv, y = "Y", x = "R", time = "t", id = "id", t_rd = 3,
                       bwselect = "cct", scheme = "pv")
r_pv_explicit$scheme
r_pv_explicit$estimates["Conventional", c("se", "ci_l", "ci_u")]
```

```{r, include = FALSE}
stopifnot(
  abs(r_cs$estimates["Conventional", "est"] - 1) < 4 * r_cs$estimates["Conventional", "se"],
  abs(r_pc$estimates["Conventional", "est"] - 1) < 4 * r_pc$estimates["Conventional", "se"],
  abs(r_pv$estimates["Conventional", "est"] - 1) < 4 * r_pv$estimates["Conventional", "se"],
  identical(c(r_cs$scheme, r_pc$scheme, r_pv$scheme), c("cs", "pc", "pv"))
)
```
